### 2025-03-14 literature.qmd
initialise

1) Knight, J., Shi, L., & Haiyuan, W. (2022). Why has China’s Inequality of Household Wealth Risen Rapidly in the Twenty‐First Century?. Review of Income and Wealth, 68(1), 109-138.

This paper tries to explain the rise in wealth inequality in 21st century China. It used the CHIP data in 2002 and 2013, emphasizing that these two releases are rigorously comparable. It attempts to explain the rise in terms of differential saving rate, differential house price inflation, income from wealth and a growing urban-rural wealth disparity. It found that the high saving rate of the rich, the housing boom, especially for the already expensive houses, real capital gains from houses, and urban-rural wealth disparity are the main causes for the rising wealth inequality, while the impact of income from wealth excluding real capital gain is marginal. This is a very important citation since it contains many technical details for wealth inequality studies. * Since housing value is missing in CHIP and housing is over 50% of net wealth for Chinese households, I abandoned using it and pivoted to the CHFS data (China Household Finance Program by a big university) instead.

2) Zhang, M. L., Brooks, M. R., Ding, D., Ding, H., He, H., Lu, J., & Mano, R. (2018). China’s high savings: drivers, prospects, and policies. International Monetary Fund.

This paper provides a comprehensive empirical review of China’s saving rates across sectors. It offers some account of historical events and their impact on saving rates. It briefly talks about the relation between income inequality and savings inequality. It mentions some fancy terms that I could reference in my paper, such as financial repression, which could be related to income and wealth inequality. It also has some regression examples with Chinese characteristics that I can learn from, such as using provincial dummies and SOE dummies. In sum, this paper provides a good example of how to construct historical narratives for empirical analysis.

3) Piketty, T., Yang, L., & Zucman, G. (2019). Capital accumulation, private property, and rising inequality in China, 1978–2015. American Economic Review, 109(7), 2469-2496.

This paper utilised various data sources to estimate the state of inequality in China. It constructed the asset side of the national balance sheet of China from 1978 to 2015, which shows the growth and composition of national wealth. It has the Piketty-style description of inequality comparing the income and wealth equality across groups. It has data for the ownership structure of China’s national wealth and corporations. There are rich comparisons between China and other countries. I could cite the data directly from this paper or reproduce certain graphs using the CHFS data. Since this paper uses various sources of data, I could cite the results in this paper to confirm or confront the results in my paper. This paper is an excellent source of background information for the literature review section of my paper.

4) Kanbur, R., Wang, Y., & Zhang, X. (2021). The great Chinese inequality turnaround. Journal of Comparative Economics, 49(2), 467-482.

This paper counters the popular view that China’s inequality has still been rising in recent years and argues that China’s inequality has plateaued. This paper used CHIP and CFPS data. It has detailed documentation and decomposition of Gini index and generalised entropy indexes, and it shows that China’s inequality peaked around 2010. It has an interesting decomposition of Theil index into within-group and between-group components. The decomposition shows that much of the Theil’s T is contributed by the urban-rural divide, which significantly shrank after 2010. The coastal-inland divide was small before the reforms, but it also contributed almost 0.05 to the overall Theil’s T in the mid 2000s, which has been declining since then. The results are counter-intuitive to me as the within urban-rural contribution is tiny compared to the between-urban-rural contribution. Another counter-intuitive observation is that the authors use two different sources and get vastly different Gini and Theil indexes. Regarding my paper, I can decompose within-province and between-province. The authors attributed the moderation of inequality to government policies. The within-between group decomposition seems to be novel and easy to implement. This paper talks about the Lewis turning point and the Kuznets turning point, which I am not familiar with.

5) Schettino, F., Gabriele, A., & Khan, H. A. (2021). Polarization and the middle class in China: A non-parametric evaluation using CHNS and CHIP data. Structural Change and Economic Dynamics, 57, 251-264.

This paper studies the polarization of China and its impact on the middle class. Polarization is an interesting intersectional idea of economics and sociology. The idea suggests that with rising polarization, the alienation between groups in a society worsens, which might cause social conflicts. This paper used Relative Distribution to distinguish between pure growth effects and pure distributional effects, and examine how the two effects have influenced the structure and size of the national middle class. This paper found that between 2000 and 2015, rapid economic growth of China lifted most segments of the population (location effect), and led to a rise in households joining the “global middle class”, but the “shape effects” or the “hollowing out” of the middle-income groups occurred. Interesting way of decomposing distribution changes into pure growth and pure distribution. I can cite their results or replicate their results with the new 2018 data. I have no idea why they did not use the 2018 data, given they wrote this in 2021.

6) Zhang, J. (2021). A survey on income inequality in China. Journal of Economic Literature, 59(4), 1191-1239.

This is a comprehensive review of works on China’s inequality. It’s a good resource to look for citations. Its findings resemble 4) that inequality in China has been rising until the 2010s when the trend slightly reversed. The trends of urban-rural disparity and regional inequality are closely correlated with the overall inequality from the 1980s to the 2000s, but no longer after the 2010s. It reported a negative correlation between inequality and intergenerational mobility, consistent with the Great Gatsby Curve in developed countries.

7) Mookherjee, D., & Shorrocks, A. (1982). A decomposition analysis of the trend in UK income inequality. The Economic Journal, 92(368), 886-902.

This is a classical piece on how to decompose inequality into within-group and between-group effects. It also has examples on how to present the decomposed effects in a chronological way. My paper might reference this paper for methodologies.

8) Liu, Y., Hao, Y., & Lu, Z. N. (2022). Health shock, medical insurance and financial asset allocation: evidence from CHFS in China. Health Economics Review, 12(1), 52.

Since I have decided no longer to use the CHIP data, I need to find some examples of using CHFS data. This is a study of how various types of health insurance help households deal with health shocks using the CHFS data. It also talks about the relation between health insurance and financial mobility. It has some examples of dividing assets into safe and risky assets. I could reference this paper for technical details.

9) Zeng, T., & Zhu, S. (2022). The mobility of top earnings, income, and wealth in China: Facts from the 2011–2017 China household finance survey. Journal of Asian Economics, 80, 101461.

This is a paper that uses the CHFS data to study wealth inequality and inter-generational mobility. It has many examples of presenting income and wealth compositions of different income groups. It noticed that the top 1% allocated a much higher proportion of wealth to private equity, which also fluctuates a lot. I might check if my results are similar to theirs if I also start from the year 2011.

10) Tan, J., Zeng, T., Zhu, S., D’ambrosio, C., Gan, L., He, H., & Zhang, J. (2017). Earnings, income, and wealth distributions in China: Facts from the 2011 China Household Finance Survey. Working Paper.

Another example of using the CHFS data to study inequality. This paper found that the level of inequality in 2011 was comparable to that of the U.S. in 1928. It has many graphs and technical details. I can use it in the literature review and in the data part.

11) Gan, L., Yin, Z., & Tan, J. (2016). Report on the development of household finance in rural China (2014). Springer Singapore.

This is a paper written by the project team of CHFS. It has a very detailed introduction of the data, including information such as the refusal rate, household composition, and gender structure. It then went on with a highly detailed report of the development of household finance in rural China. I will cite this paper for technical details of the data and some statistics.

12) Xie, Y., & Jin, Y. (2015). Household wealth in China. Chinese sociological review, 47(3), 203-229.

Another more cited empirical paper on China’s wealth inequality but it did not use CHFS. I can reference it for facts in literature review. These last three papers are basically presenting data and have less to do with inequality measure decomposition. The CHFS dataset is huge with thousands of variables, and thus it is important to follow how they aggregate individual asset / liability items into broad asset classes.

### 2025-04-03 first_draft/draft.qmd
update

---
title: "Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS"
format:
    pdf:
        documentclass: article 
        author: Chenning Xu
        number-sections: true
        margin-top: 1in
        margin-bottom: 1in
        margin-left: 1in
        margin-right: 1in
        linestretch:  1.2
        fontsize: 10pt
    html: default    
bibliography: tom.bib
---

# Introduction
China has experienced remarkable economic growth in recent decades following its reform and opening-up policy initiated in 1978. However, this period has also been marked by a significant rise in income and wealth inequality [@piketty_capital_2019]. Some scholars have provided evidence suggesting that China's overall inequality peaked around 2010 and subsequently began to decline [@kanbur_great_2021; @zhang_survey_2021]. It is also well-documented that China's inequality is strongly intertwined with its urban-rural and coastal-inland divisions [@piketty_capital_2019; @zhang_survey_2021]. This paper utilizes data from the 2017 and 2019 China Household Finance Survey (CHFS) to examine the current state of China's wealth inequality. Specifically, it investigates whether the latest publicly available data supports the view that China's inequality has continued to decline since around 2010, as well as how China's wealth inequality relates to urban-rural and regional disparities.

# Background
This section presents some notable findings on China's income and wealth inequality in the existing literature.

China has experienced a considerable increase in income and wealth inequality since the reform and opening-up policy initiated in 1978. Using diverse sources, including tax records, household surveys, and private wealth rankings, @piketty_capital_2019 provide a detailed analysis of the transformation in China's inequality profile. Their findings show that between 1978 and 2015, the income share of the top 10% increased from 27% to 41%, while the share of the bottom 50% decreased from 27% to 15%. Wealth concentration also intensified significantly: the wealth share held by the top 10% rose from around 40% to nearly 70%, whereas the bottom 50%'s share dropped from over 15% to about 5%. Additionally, they documented an increase in the urban-to-rural per capita income ratio, from less than 200% in 1978 to approximately 350% by 2015. Humorously noting this stark shift, they remarked that China had transitioned from Nordic-level equality to U.S.-level inequality during the reform era.

@knight_why_2022 further document a rapid rise in China's household wealth inequality between 2002 and 2013, reporting that the national Gini coefficient for household wealth per capita increased markedly from 0.50 to 0.62. They identify housing wealth as the primary driver behind this increase, with its share rising from 53% to 73% of total household wealth, and its contribution to overall wealth inequality growing from 64% to 79%. The authors emphasize that differential savings rates across income groups, uneven house price inflation—particularly pronounced in major urban centers—and widening urban-rural disparities significantly intensified wealth inequality during this period.

The income and wealth inequality highly intersects with urban-rural and reginal disparities. @zhang_survey_2021 highlights significant rural-urban and regional disparities underlying China's rising inequality since its economic reforms. The urban-rural income ratio increased sharply, peaking at around 3.3 in 2009 before declining slightly to 2.7 by 2019, still substantially higher than pre-reform levels. Despite a notable decline in overall inequality after 2010—with the national Gini coefficient decreasing from a peak of 0.491 in 2008 to approximately 0.465 in 2019—regional inequality remains pronounced. For instance, per capita GDP in the eastern coastal region was roughly 2.6 times higher than in the western region by 2018. Additionally, rural-to-urban migration grew dramatically from fewer than 20 million in 1990 to approximately 290 million by 2019, reflecting both the economic disparity between regions and the persistent urban-rural divide, exacerbated by restricted access to urban public services under the Hukou household registration system.

Consistent with the finding in @zhang_survey_2021 that China's inequality peaked in 2008 and 2009, @kanbur_great_2021 similarly document that China's inequality increased sharply after economic reforms but started to plateau around 2010. They find the national Gini coefficient peaked at about 0.525 in 2010, subsequently declining to approximately 0.476 in 2016 before rising slightly again. Consistent with Zhang (2021), they attribute this recent turnaround primarily to narrowing rural-urban and coastal-inland disparities, driven by demographic changes, tightening rural labor markets, and rising rural wages.

# Data
Using the relatively new and underexplored 2017 and 2019 China Household Finance Survey (CHFS) datasets, this paper aims to compare its findings with results from earlier literature. CHFS is a national household survey initiated by the Southwestern University of Finance and Economics (SWUFE) in 2009, with the first results published in 2011. It focuses specifically on collecting detailed, micro-level information on household finances (@kanbur_great_2021). Since 2017, CHFS has made available aggregated household-level data on wealth variables such as housing values, though the original micro-level data remains proprietary. The availability of these aggregated data allows this paper to analyze China's wealth inequality in 2017 and 2019, comparing these recent results with earlier analyses by the CHFS research team and other scholars who utilized different data sources.

The household wealth variables consistently available in both 2017 and 2019 include total household assets and debts, financial assets and debts, agricultural assets and debts, business assets and debts, land, housing assets and debts, commercial property assets and debts, vehicle assets and debts, education debts, credit debts, medical debts, and other miscellaneous assets and debts. In 2019, an additional variable, garage assets, was reported separately from housing assets; this variable was not present in 2017.

For consistency, this paper calculates equity values by deducting debts from corresponding assets, except for land, which is typically fully owned and non-transferable. Due to its relatively small aggregate value, garage assets in 2019 are recombined with housing assets for analytical consistency.

Given the relatively short period between 2017 and 2019, this study aims to identify stable patterns in China's wealth inequality using data from these two years. Additionally, it compares the findings with earlier studies that utilize the same CHFS dataset to infer recent trends in wealth inequality[@tan_earnings_2017].

# Analysis
## Change of Gini from 2011 to 2019
This section presents the Gini index analysis using the 2017 and 2019 CHFS data. The statistics from the 2011 CHFS data is from @tan_earnings_2017, which is work from CFHS's research team. All the income and wealth statistics are aggregated at the household level.

| Statistic | Income 2011 | Income 2017 | Income 2019 | Wealth 2011 | Wealth 2017 | Wealth 2019 |
|-----------|-------------|-------------|-------------|-------------|-------------|-------------|
| Mean      | 60,053     | 91,324     | 73,921     | 687,061    | 909,283    | 860,754    |
| Median    | 28,312     | 57,142     | 42,279     | 197,150    | 355,837    | 312,949    |
| Gini      | 0.664      | 0.592      | 0.637      | 0.761      | 0.693      | 0.705      |
| N         | 8,438      | 38,994     | 33,621     | 8,438      | 38,994     | 33,621     |

Compared to the 2011 data, the Gini coefficient for both income and wealth has decreased in 2017 and 2019, which is consistent with the findings in @kanbur_great_2021 and @zhang_survey_2021. Both income and wealth are higher by mean and median in 2017 than 2019, which is a strange finding that is likely due to a change in the survey design. The Gini coefficients for income and wealth are slighly higher in 2019 than 2017, but the difference is not significant enough to draw a conclusion espcially given the change in survey design.

## Factor decomposition of wealth inequality

| 2017 Wealth      | Gini Correlation | Gini Coefficient | Share   | Contribution |
|-------------------|------------------|------------------|---------|--------------|
| Houesing Equity  | 0.9661          | 0.7181          | 0.7204  | 0.7211       |
| Commercial Equity | 0.8153          | 0.9867          | 0.0202  | 0.0234       |
| Car Equity        | 0.6537          | 0.9099          | 0.0304  | 0.0261       |
| Financial Equity  | 0.7943          | 0.7915          | 0.1228  | 0.1114       |
| Other Equity      | 0.6354          | 0.9332          | 0.0198  | 0.0170       |
| Educational Debt  | -0.3040         | 0.9875          | 0.0011  | -0.0005      |
| Credit Debt       | 0.3427          | 0.9903          | 0.0013  | 0.0007       |
| Medical Debt      | -0.5763         | 0.9816          | 0.0019  | -0.0015      |

| 2019 Wealth      | Gini Correlation | Gini Coefficient | Share   | Contribution |
|-------------------|------------------|------------------|---------|--------------|
| Housing Equity  | 0.9603          | 0.7362          | 0.6798  | 0.6817       |
| Commercial Equity | 0.8052          | 0.9920          | 0.0133  | 0.0151       |
| Car Equity        | 0.6818          | 0.8569          | 0.0306  | 0.0254       |
| Financial Equity  | 0.8031          | 0.7883          | 0.1386  | 0.1245       |
| Other Equity      | 0.7336          | 1.0215          | 0.0237  | 0.0251       |
| Educational Debt  | -0.3742         | 0.9889          | 0.0009  | -0.0005      |
| Credit Debt       | 0.3613          | 0.9910          | 0.0009  | 0.0004       |
| Medical Debt      | -0.5068         | 0.9824          | 0.0022  | -0.0015      |

The factor decompostion of the Gini coefficient for household wealth in two years shows that housing equity is still the largest and dominant contributor to China's wealth inequality, contributing around 70% to the overall wealth Gini. Financial equity follows by contributing to 11.1% in 2017 and 12.5% in 2019 to the overall wealth Gini. Commercial equity such as agricultural and business assets contrbutes to a smaller portion, by about 5% in 2017 and 4% in 2019.

# References

### 2025-04-03 first_draft/ref.qmd
update

--- 2025-03-14 literature.qmd
@@
 
-7) Mookherjee, D., & Shorrocks, A. (1982). A decomposition analysis of the trend in UK income inequality. The Economic Journal, 92(368), 886-902.
-
-This is a classical piece on how to decompose inequality into within-group and between-group effects.
-It also has examples on how to present the decomposed effects in a chronological way.
-My paper might reference this paper for methodologies.
-
 8) Liu, Y., Hao, Y., & Lu, Z. N. (2022). Health shock, medical insurance and financial asset allocation: evidence from CHFS in China. Health Economics Review, 12(1), 52.

### 2025-04-04 first_draft/draft.qmd
update

@@
 It focuses specifically on collecting detailed, micro-level information on household finances (@kanbur_great_2021).
-Since 2017, CHFS has made available aggregated household-level data on wealth variables such as housing values, though the original micro-level data remains proprietary.
-The availability of these aggregated data allows this paper to analyze China's wealth inequality in 2017 and 2019, comparing these recent results with earlier analyses by the CHFS research team and other scholars who utilized different data sources.
 
@@
 |-----------|-------------|-------------|-------------|-------------|-------------|-------------|
-| Mean      | 60,053     | 91,324     | 73,921     | 687,061    | 909,283    | 860,754    |
-| Median    | 28,312     | 57,142     | 42,279     | 197,150    | 355,837    | 312,949    |
-| Gini      | 0.664      | 0.592      | 0.637      | 0.761      | 0.693      | 0.705      |
-| N         | 8,438      | 38,994     | 33,621     | 8,438      | 38,994     | 33,621     |
+| Mean      | 60,053     | 89,634     | 94,005     | 687,061    | 839,008    | 1,085,589    |
+| Median    | 28,312     | 54,108     | 57,035     | 197,150    | 322,515    | 457,600    |
+| Gini      | 0.664      | 0.605      | 0.614     | 0.761      | 0.701      | 0.667      |
+| N         | 8,438      | 38,994     | 33,387     | 8,438      | 38,994     | 33,387     |
 
-Compared to the 2011 data, the Gini coefficient for both income and wealth has decreased in 2017 and 2019, which is consistent with the findings in @kanbur_great_2021 and @zhang_survey_2021.
-Both income and wealth are higher by mean and median in 2017 than 2019, which is a strange finding that is likely due to a change in the survey design.
-The Gini coefficients for income and wealth are slighly higher in 2019 than 2017, but the difference is not significant enough to draw a conclusion espcially given the change in survey design.
+Compared to the 2011 data, the Gini coefficient for income has decresed from 0.664 in 2011 to 0.605 in 2017 and remained stable in 2019.
+The Gini coefficient for wealth has decreased from 0.761 in 2011 to 0.701 in 2017 and continued to decrease to 0.667 in 2019.
+The results show that China's income and wealth inequality has been decreasing since 2011.
+The finding is consistent with the findings in @kanbur_great_2021 and @zhang_survey_2021, that China's inequality peaked around 2010 and started to decrease since then.
 
@@
 |-------------------|------------------|------------------|---------|--------------|
-| Houesing Equity  | 0.9661          | 0.7181          | 0.7204  | 0.7211       |
-| Commercial Equity | 0.8153          | 0.9867          | 0.0202  | 0.0234       |
-| Car Equity        | 0.6537          | 0.9099          | 0.0304  | 0.0261       |
-| Financial Equity  | 0.7943          | 0.7915          | 0.1228  | 0.1114       |
-| Other Equity      | 0.6354          | 0.9332          | 0.0198  | 0.0170       |
-| Educational Debt  | -0.3040         | 0.9875          | 0.0011  | -0.0005      |
-| Credit Debt       | 0.3427          | 0.9903          | 0.0013  | 0.0007       |
-| Medical Debt      | -0.5763         | 0.9816          | 0.0019  | -0.0015      |
+| Housing Equity  | 0.9610          | 0.7244          | 0.6996  | 0.6950       |
+| Land Asset      | 0.4722          | 0.9204          | 0.0392  | 0.0243       |
+| Commercial Equity | 0.8246          | 1.1989          | 0.0667  | 0.0941       |
+| Commercial Property | 0.8290          | 0.9890          | 0.0213  | 0.0250       |
+| Car Equity        | 0.6648          | 0.9037          | 0.0339  | 0.0290       |
+| Financial Equity  | 0.8022          | 0.7964          | 0.1232  | 0.1123       |
+| Other Equity      | 0.6485          | 0.9438          | 0.0213  | 0.0186       |
+| Educational Debt  | -0.3166         | 0.9846          | 0.0014  | -0.0006      |
+| Credit Card Debt  | 0.3678          | 0.9899          | 0.0015  | 0.0008       |
+| Medical Debt      | -0.5801         | 0.9786          | 0.0023  | -0.0018       |
 
@@
 |-------------------|------------------|------------------|---------|--------------|
-| Housing Equity  | 0.9603          | 0.7362          | 0.6798  | 0.6817       |
-| Commercial Equity | 0.8052          | 0.9920          | 0.0133  | 0.0151       |
-| Car Equity        | 0.6818          | 0.8569          | 0.0306  | 0.0254       |
-| Financial Equity  | 0.8031          | 0.7883          | 0.1386  | 0.1245       |
-| Other Equity      | 0.7336          | 1.0215          | 0.0237  | 0.0251       |
-| Educational Debt  | -0.3742         | 0.9889          | 0.0009  | -0.0005      |
-| Credit Debt       | 0.3613          | 0.9910          | 0.0009  | 0.0004       |
-| Medical Debt      | -0.5068         | 0.9824          | 0.0022  | -0.0015      |
+| Housing Equity  | 0.9591          | 0.7015          | 0.6834  | 0.6876       |
+| Land Asset      | 0.7155          | 0.9700          | 0.0593  | 0.0615       |
+| Commercial Equity | 0.6812          | 1.2014          | 0.0359  | 0.0440       |
+| Commercial Property | 0.7778          | 0.9886          | 0.0144  | 0.0166       |
+| Car Equity        | 0.6651          | 0.8224          | 0.0344  | 0.0281       |
+| Financial Equity  | 0.7895          | 0.7484          | 0.1532  | 0.1354       |
+| Other Equity      | 0.7171          | 1.0296          | 0.0224  | 0.0247       |
+| Educational Debt  | -0.4199         | 0.9885          | 0.0008  | -0.0005      |
+| Credit Card Debt  | 0.3029          | 0.9854          | 0.0012  | 0.0005       |
+| Medical Debt      | -0.5591         | 0.9844          | 0.0016  | -0.0014       |
 
@@
 Financial equity follows by contributing to 11.1% in 2017 and 12.5% in 2019 to the overall wealth Gini.
-Commercial equity such as agricultural and business assets contrbutes to a smaller portion, by about 5% in 2017 and 4% in 2019.
+Commercial equity such as agricultural and business assets contrbutes to a relatively large portion in 2017 but to a smaller portion in 2019.
+Land asset contributes to a relatively large portion in 2017 but to a smaller portion in 2019.
+The cause of this is uninvestigated, which could be due to the change in survey design.
+Commercial property and car equity contribute to a relatively small portion in both years.
+Educational debt and medical debt negatively correlates with wealth inequality, which shows that the poor is disporportionately burdened by educational and medical debts, although their contribution to the overall wealth Gini is small.
+
+This high contribution of housing equity to wealth inequality is consistent with other studies.
+@xie_household_2015 used the 2012 CHPS data and documented that housing equity conprises of 71.6% of household wealth in China.
+@kanbur_great_2021 used the 2002 and 2013 CHIP data and found that housing euqity contribues to 64% of wealth inequality in 2002 and 79% in 2013.
+@tan_earnings_2017 documented with the 2011 CHFS data that the housing asset to total asset ratio was 66.73%, and the same ratio was 71.08% in 2017 and 67.23% in 2019.
+It appears that housing remains the largest wealth holding for Chinese households in 2017 and 2019, and housing is still the largest contributor to wealth inequality with no clear sign of change.
+
+## Group decomposition of wealth inequality by rural and urban
+This section presents the group decomposition of wealth inequality by the urban-rural and regional divisions.
+
+| Group | Population Share | Mean | Relative Mean | Median Income | Wealth Share | Gini |
+|-------|-----------------|------|---------------|---------------|--------------|------|
+| 2017 Urban | 0.6235 | 1,146,510 | 1.3665 | 66,760 | 0.8521 | 0.6573 |
+| 2017 Rural | 0.3765 | 329,715 | 0.3930 | 23,660 | 0.1479 | 0.6945 |
+| 2019 Urban | 0.6500 | 1,461,675 | 1.3464 | 69,600 | 0.8752 | 0.6165 |
+| 2019 Rural | 0.3500 | 387,130 | 0.3566 | 27,541 | 0.1248 | 0.6629 |
+
+The summary statistics above shows that in both 2017 and 2019, the mean and median income levels are significantly higher in urban areas compared to rural areas.
+The mean wealth in urban areas is 3.48 times the mean wealth in rural areas in 2017, and 3.78 times in 2019.
+The median wealth in urban areas is 2.82 times the median wealth in rural areas in 2017, and 2.52 times in 2019.
+
+Notably, the Gini coefficient for wealth is higher in rural areas than in urban areas in both 2017 and 2019.
+
+The decomposition of the Gini coefficient into within-group, between-group, and residual components shows[@lambert_inequality_1993]:
+
+| Year | Total Gini | Within-group | Between-group | Residual |
+|------|------------|--------------|---------------|----------|
+| 2017 | 100.00 | 55.35 | 32.61 | 12.04 |
+| 2019 | 100.00 | 56.79 | 33.68 | 9.53 |
+
+This decomposition reveals that about 55-57% of total wealth inequality is due to inequality within urban and rural groups, while about 33-34% is due to inequality between urban and rural areas.
+The residual component, representing the interaction between the within and between components, accounts for about 10-12% of total inequality.
+The stability of these components between 2017 and 2019 suggests that the urban-rural parity remains a significant driver of wealth inequality in China.
+
+## Group decomposition of wealth inequality by region
 

### 2025-04-05 first_draft/draft.qmd
updtae

@@
 
+Notably, the growth rate of mean and median income from 2011 to 2017 is clearly higher than that of wealth but from 2017 to 2019, the growth of mean and median wealth surpasses that of income by a great margin.
+
 ## Factor decomposition of wealth inequality
@@
 ## Group decomposition of wealth inequality by rural and urban
-This section presents the group decomposition of wealth inequality by the urban-rural and regional divisions.
+The following two sections present the group decomposition of wealth inequality by the urban-rural and regional divisions.
 
-| Group | Population Share | Mean | Relative Mean | Median Income | Wealth Share | Gini |
-|-------|-----------------|------|---------------|---------------|--------------|------|
+| Group      | Sample Share | Mean    | Relative Mean | Median Income | Wealth Share | Gini |
+|------------|--------------|---------|---------------|---------------|--------------|------|
 | 2017 Urban | 0.6235 | 1,146,510 | 1.3665 | 66,760 | 0.8521 | 0.6573 |
@@
 
-Notably, the Gini coefficient for wealth is higher in rural areas than in urban areas in both 2017 and 2019.
+Notably, the Gini coefficient for wealth is higher in rural areas than in urban areas in both 2017 and 2019. The Gini coefficient for both urban rural areas declined from 2017 to 2019.
 
@@
 
+| Region         | Sample Share | Mean        | Relative Mean | Median Income | Wealth Share | Gini   |
+|----------------|--------------|-------------|---------------|---------------|--------------|--------|
+| 2017 East     | 0.3754       | 1,391,421   | 1.6584        | 556,993       | 0.6226       | 0.6747 |
+| 2017 Middle   | 0.2659       | 550,976     | 0.6567        | 272,235       | 0.1746       | 0.6648 |
+| 2017 West     | 0.2628       | 496,937     | 0.5923        | 252,341       | 0.1556       | 0.6569 |
+| 2017 Northeast| 0.0960       | 412,387     | 0.4915        | 212,512       | 0.0472       | 0.6513 |
+| 2019 East     | 0.3987       | 1,658,323   | 1.5276        | 760,140       | 0.6090       | 0.6305 |
+| 2019 Middle   | 0.2651       | 677,697     | 0.6243        | 372,434       | 0.1655       | 0.5997 |
+| 2019 West     | 0.2400       | 799,535     | 0.7365        | 343,542       | 0.1768       | 0.6947 |
+| 2019 Northeast| 0.0962       | 549,715     | 0.5064        | 291,085       | 0.0487       | 0.6369 |
+
+The above summary statistics shows that the mean and median wealth in the East is significantly higher than in the other regions in both 2017 and 2019.
+Notably, the mean and median wealth of the West-area households surpassed that of the Middle in 2019, but West's Gini coefficient rose to the highest among all regions in 2019 as well.
+
+The decomposition of the Gini coefficient into within-group, between-group, and residual components shows:
+
+| Year | Total Gini | Within-group | Between-group | Residual |
+|------|------------|--------------|---------------|----------|
+| 2017 | 100.00 | 31.16 | 36.88 | 31.96 |
+| 2019 | 100.00 | 31.69 | 33.77 | 34.54 |
+
+The group decompsition shows that similar to the urban-rural decomposition, the between-group component contribues over 30% to the overall wealth inequality in both 2017 and 2019.
+The between group contribution saw a slight decrease from 2017 to 2019 with a similari increase in the residual component.
+This might signal some modest convergence in household wealth between regions.
+
 # References

### 2025-04-20 report/draft.qmd
update

--- 2025-04-05 first_draft/draft.qmd
@@
         documentclass: article 
-        author: Chenning Xu
         number-sections: true
@@
         fontsize: 10pt
-    html: default    
+        fig-pos: "H"
+        author: "Chenning Xu"
 bibliography: tom.bib
@@
 
-| Statistic | Income 2011 | Income 2017 | Income 2019 | Wealth 2011 | Wealth 2017 | Wealth 2019 |
-|-----------|-------------|-------------|-------------|-------------|-------------|-------------|
-| Mean      | 60,053     | 89,634     | 94,005     | 687,061    | 839,008    | 1,085,589    |
-| Median    | 28,312     | 54,108     | 57,035     | 197,150    | 322,515    | 457,600    |
-| Gini      | 0.664      | 0.605      | 0.614     | 0.761      | 0.701      | 0.667      |
-| N         | 8,438      | 38,994     | 33,387     | 8,438      | 38,994     | 33,387     |
+::: {#tbl-tbl1}
+
+{{< include summary_stats.txt >}}
+
+Source: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.
+
+Mean, Median, and Gini Coefficient for Income and Wealth in China, 2011, 2017, and 2019
+:::
 
@@
 
-| 2017 Wealth      | Gini Correlation | Gini Coefficient | Share   | Contribution |
-|-------------------|------------------|------------------|---------|--------------|
-| Housing Equity  | 0.9610          | 0.7244          | 0.6996  | 0.6950       |
-| Land Asset      | 0.4722          | 0.9204          | 0.0392  | 0.0243       |
-| Commercial Equity | 0.8246          | 1.1989          | 0.0667  | 0.0941       |
-| Commercial Property | 0.8290          | 0.9890          | 0.0213  | 0.0250       |
-| Car Equity        | 0.6648          | 0.9037          | 0.0339  | 0.0290       |
-| Financial Equity  | 0.8022          | 0.7964          | 0.1232  | 0.1123       |
-| Other Equity      | 0.6485          | 0.9438          | 0.0213  | 0.0186       |
-| Educational Debt  | -0.3166         | 0.9846          | 0.0014  | -0.0006      |
-| Credit Card Debt  | 0.3678          | 0.9899          | 0.0015  | 0.0008       |
-| Medical Debt      | -0.5801         | 0.9786          | 0.0023  | -0.0018       |
+::: {#tbl-tbl2}
 
-| 2019 Wealth      | Gini Correlation | Gini Coefficient | Share   | Contribution |
-|-------------------|------------------|------------------|---------|--------------|
-| Housing Equity  | 0.9591          | 0.7015          | 0.6834  | 0.6876       |
-| Land Asset      | 0.7155          | 0.9700          | 0.0593  | 0.0615       |
-| Commercial Equity | 0.6812          | 1.2014          | 0.0359  | 0.0440       |
-| Commercial Property | 0.7778          | 0.9886          | 0.0144  | 0.0166       |
-| Car Equity        | 0.6651          | 0.8224          | 0.0344  | 0.0281       |
-| Financial Equity  | 0.7895          | 0.7484          | 0.1532  | 0.1354       |
-| Other Equity      | 0.7171          | 1.0296          | 0.0224  | 0.0247       |
-| Educational Debt  | -0.4199         | 0.9885          | 0.0008  | -0.0005      |
-| Credit Card Debt  | 0.3029          | 0.9854          | 0.0012  | 0.0005       |
-| Medical Debt      | -0.5591         | 0.9844          | 0.0016  | -0.0014       |
+{{< include 2017_gini_decom.txt >}}
+
+Source: Author's calculation based on the 2017 CHFS data.
+
+Decomposition of Wealth Gini Coefficient by Urban-Rural Division, 2017
+:::
+
+::: {#tbl-tbl3}
+
+{{< include 2019_gini_decom.txt >}}
+
+Source: Author's calculation based on the 2019 CHFS data.
+
+Decomposition of Wealth Gini Coefficient by Urban-Rural Division, 2019
+:::
 
@@
 
-| Group      | Sample Share | Mean    | Relative Mean | Median Income | Wealth Share | Gini |
-|------------|--------------|---------|---------------|---------------|--------------|------|
-| 2017 Urban | 0.6235 | 1,146,510 | 1.3665 | 66,760 | 0.8521 | 0.6573 |
-| 2017 Rural | 0.3765 | 329,715 | 0.3930 | 23,660 | 0.1479 | 0.6945 |
-| 2019 Urban | 0.6500 | 1,461,675 | 1.3464 | 69,600 | 0.8752 | 0.6165 |
-| 2019 Rural | 0.3500 | 387,130 | 0.3566 | 27,541 | 0.1248 | 0.6629 |
+::: {#tbl-tbl4}
+
+{{< include rural_urban_summary.txt >}}
+
+Source: Author's calculation based on the 2017 and 2019 CHFS data.
+
+Summary Statistics of Urban and Rural Income and Wealth, 2017 and 2019
+:::
 
@@
 
-| Year | Total Gini | Within-group | Between-group | Residual |
-|------|------------|--------------|---------------|----------|
-| 2017 | 100.00 | 55.35 | 32.61 | 12.04 |
-| 2019 | 100.00 | 56.79 | 33.68 | 9.53 |
+::: {#tbl-urbanrural-decomp}
+
+{{< include urban_rural_gini_decomp.txt >}}
+
+Source: Author's calculation based on the 2017 and 2019 CHFS data.
+
+Decomposition of Wealth Gini Coefficient by Urban-Rural Division, 2017 and 2019 (%)
+:::
 
@@
 
-| Region         | Sample Share | Mean        | Relative Mean | Median Income | Wealth Share | Gini   |
-|----------------|--------------|-------------|---------------|---------------|--------------|--------|
-| 2017 East     | 0.3754       | 1,391,421   | 1.6584        | 556,993       | 0.6226       | 0.6747 |
-| 2017 Middle   | 0.2659       | 550,976     | 0.6567        | 272,235       | 0.1746       | 0.6648 |
-| 2017 West     | 0.2628       | 496,937     | 0.5923        | 252,341       | 0.1556       | 0.6569 |
-| 2017 Northeast| 0.0960       | 412,387     | 0.4915        | 212,512       | 0.0472       | 0.6513 |
-| 2019 East     | 0.3987       | 1,658,323   | 1.5276        | 760,140       | 0.6090       | 0.6305 |
-| 2019 Middle   | 0.2651       | 677,697     | 0.6243        | 372,434       | 0.1655       | 0.5997 |
-| 2019 West     | 0.2400       | 799,535     | 0.7365        | 343,542       | 0.1768       | 0.6947 |
-| 2019 Northeast| 0.0962       | 549,715     | 0.5064        | 291,085       | 0.0487       | 0.6369 |
+::: {#tbl-region-summary}
+
+{{< include region_summary.txt >}}
+
+Source: Author's calculation based on the 2017 and 2019 CHFS data.
+
+Summary Statistics by Region, 2017 and 2019
+:::
 
@@
 
-The decomposition of the Gini coefficient into within-group, between-group, and residual components shows:
+::: {#tbl-region-decomp}
 
-| Year | Total Gini | Within-group | Between-group | Residual |
-|------|------------|--------------|---------------|----------|
-| 2017 | 100.00 | 31.16 | 36.88 | 31.96 |
-| 2019 | 100.00 | 31.69 | 33.77 | 34.54 |
+{{< include region_gini_decomp.txt >}}
+
+Source: Author's calculation based on the 2017 and 2019 CHFS data.
+
+Decomposition of Wealth Gini Coefficient by Region, 2017 and 2019 (%)
+:::
 
@@
 
+## What's next
+I will read more papers and contrast the results with data in this study. I am not sure about what else I can do for this paper. I would like to hear your suggestions. Thanks.
+
 # References

### 2025-04-22 report/draft.qmd
update

@@
 
-Source: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.
+All figures are reported in RMB at constant 2017 prices.
+
+Sources: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.
 
@@
 ## Factor decomposition of wealth inequality
+This section examines the composition and distribution of household wealth in China by analyzing specific asset and debt categories using data from the 2017 and 2019 China Household Finance Survey (CHFS).
+This detailed decomposition provides insights into the relative importance of different components contributing to wealth inequality.
+
+::: {#tbl-wealth-summary}
+
+{{< include wealth_summary.txt >}}
+
+All figures are reported in RMB at constant 2017 prices.
+
+Source: Author's calculation based on the 2017 and 2019 CHFS data.
+
+Summary Statistics of Wealth, 2017 and 2019
+:::
+
+The above table intuitively shows the state of household wealth allocation in China in 2017 and 2019.
+Clearly the mean and median value of house dominates other assets.
+It is also notable that despite the high inequality in housing assets, as it is shown in the next table, the proportion of households that have no houses at all is fairly low at just 10%.
+It is also surprising that the housing debt is very low compared to the asset value, and 82% of the households have no house debts.
 
@@
 ## Group decomposition of wealth inequality by rural and urban
-The following two sections present the group decomposition of wealth inequality by the urban-rural and regional divisions.
+Group decomposition of wealth inequality by rural and urban
+
+To analyze the sources of wealth inequality, this paper employs the classic Gini decomposition identity initially developed by Bhattacharya and Mahalanobis (1967)[@bhattacharya_regional_1967].
+The overall Gini coefficient  is decomposed as follows:
+
+$$G = G_{B} + \sum_{k} a_{k}G_{k} + R$$
+
+where $G_{B}$ is the between-group Gini, which captures inequality arising solely from differences in group means; $G_{k}$ is the within-group component, a weighted sum of the inequality within each group, with weights $a_{k}$ representing the products of each group's population share and wealth share; and $R$ is the overlap (residual) component, which captures inequality due to the intersection of group wealth distributions.
+Lambert and Aronson (1993) provide a geometric interpretation of this overlap term, demonstrating that $R$ corresponds to the Lorenz curve area generated by reranking when subgroup wealth distributions overlap.
+A positive $R$ thus indicates significant intersections among wealth distributions of different groups[@lambert_inequality_1993].
 
@@
 
-Summary Statistics of Urban and Rural Income and Wealth, 2017 and 2019
+Summary statistics of urban and rural income and wealth, 2017 & 2019
 :::
 
-The summary statistics above shows that in both 2017 and 2019, the mean and median income levels are significantly higher in urban areas compared to rural areas.
-The mean wealth in urban areas is 3.48 times the mean wealth in rural areas in 2017, and 3.78 times in 2019.
-The median wealth in urban areas is 2.82 times the median wealth in rural areas in 2017, and 2.52 times in 2019.
-
-Notably, the Gini coefficient for wealth is higher in rural areas than in urban areas in both 2017 and 2019. The Gini coefficient for both urban rural areas declined from 2017 to 2019.
-
-The decomposition of the Gini coefficient into within-group, between-group, and residual components shows[@lambert_inequality_1993]:
+The summary statistics highlight significant and persistent disparities between urban and rural China.
+Urban households consistently exhibit higher mean and median incomes and wealth compared to rural households.
+Urban mean wealth was 3.48 times greater than rural mean wealth in 2017 and widened slightly to 3.78 times in 2019.
+Median urban wealth was 2.82 times higher than the rural median in 2017, though this ratio decreased slightly to 2.52 in 2019.
+Additionally, the Gini coefficient for wealth is consistently higher in rural areas, though both rural and urban Gini coefficients declined modestly between 2017 and 2019.
 
@@
 
+All figures are reported in RMB at constant 2017 prices.
+
 Source: Author's calculation based on the 2017 and 2019 CHFS data.
 
-Decomposition of Wealth Gini Coefficient by Urban-Rural Division, 2017 and 2019 (%)
+Decomposition of wealth Gini coefficient by urban–rural division, 2017 & 2019 (%)
 :::
 
-This decomposition reveals that about 55-57% of total wealth inequality is due to inequality within urban and rural groups, while about 33-34% is due to inequality between urban and rural areas.
-The residual component, representing the interaction between the within and between components, accounts for about 10-12% of total inequality.
-The stability of these components between 2017 and 2019 suggests that the urban-rural parity remains a significant driver of wealth inequality in China.
-
-## Group decomposition of wealth inequality by region
+The decomposition results indicate that about 55–57% of total wealth inequality is attributable to within-group disparities among urban and rural households.
+The between-group component, reflecting the urban-rural wealth divide, contributes approximately 33–34%.
+The overlap component, representing inequality due to intersecting distributions of urban and rural wealth, accounts for around 9–12% of total inequality.
+The consistency of these shares between 2017 and 2019 underscores the continued significance of the urban-rural divide in shaping China's wealth inequality.
 
@@
 {{< include region_summary.txt >}}
+
+All figures are reported in RMB at constant 2017 prices.
 
@@
 
-## What's next
-I will read more papers and contrast the results with data in this study. I am not sure about what else I can do for this paper. I would like to hear your suggestions. Thanks.
-
 # References

### 2025-04-23 report/draft.qmd
update

@@
 
-The factor decompostion of the Gini coefficient for household wealth in two years shows that housing equity is still the largest and dominant contributor to China's wealth inequality, contributing around 70% to the overall wealth Gini.
-Financial equity follows by contributing to 11.1% in 2017 and 12.5% in 2019 to the overall wealth Gini.
-Commercial equity such as agricultural and business assets contrbutes to a relatively large portion in 2017 but to a smaller portion in 2019.
-Land asset contributes to a relatively large portion in 2017 but to a smaller portion in 2019.
-The cause of this is uninvestigated, which could be due to the change in survey design.
-Commercial property and car equity contribute to a relatively small portion in both years.
-Educational debt and medical debt negatively correlates with wealth inequality, which shows that the poor is disporportionately burdened by educational and medical debts, although their contribution to the overall wealth Gini is small.
-
+The factor decompostion of the Gini coefficient for household wealth in two years shows that housing asset dominates the wealth inequality in China.
+In both 2017 and 2019, the housing asset occupy over 70% of total household wealth and contribute to over 72% of the overall wealth Gini.
 This high contribution of housing equity to wealth inequality is consistent with other studies.
-@xie_household_2015 used the 2012 CHPS data and documented that housing equity conprises of 71.6% of household wealth in China.
-@kanbur_great_2021 used the 2002 and 2013 CHIP data and found that housing euqity contribues to 64% of wealth inequality in 2002 and 79% in 2013.
-@tan_earnings_2017 documented with the 2011 CHFS data that the housing asset to total asset ratio was 66.73%, and the same ratio was 71.08% in 2017 and 67.23% in 2019.
-It appears that housing remains the largest wealth holding for Chinese households in 2017 and 2019, and housing is still the largest contributor to wealth inequality with no clear sign of change.
+@tan_earnings_2017 documented with the 2011 CHFS data that the housing asset to total asset ratio was 66.73%, and the 2017 and 2019 74.2% and 73% shows an sharp increase in the housing asset to total asset ratio from 2011 to 2019.
+Meanwhile, @kanbur_great_2021 used the 2002 and 2013 CHIP data and found that housing euqity contribues to 64% of wealth inequality in 2002 and 79% in 2013.
+Albeit variation in the statistics across sources, the results in @tbl-tbl3 demonstrates that in 2017 and 2019, the housing market is still the most important driver of wealth inequality in China.
+However, surprisingly,
 

### 2025-04-23 report/draft.qmd
update

@@
 # Data
-Using the relatively new and underexplored 2017 and 2019 China Household Finance Survey (CHFS) datasets, this paper aims to compare its findings with results from earlier literature.
-CHFS is a national household survey initiated by the Southwestern University of Finance and Economics (SWUFE) in 2009, with the first results published in 2011.
-It focuses specifically on collecting detailed, micro-level information on household finances (@kanbur_great_2021).
+The analysis in this study utilizes data from the 2017 and 2019 waves of the China Household Finance Survey (CHFS), a nationally representative dataset collected by the Survey and Research Center for China Household Finance at Southwestern University of Finance and Economics [@gan_report_2014].
+The CHFS dataset includes household-level financial information on income, assets, debts, consumption, and demographic characteristics.
+Data are collected through face-to-face interviews using a multi-stage probability proportional to size (PPS) sampling method, ensuring representativeness at both national and provincial levels.
+Sample weights provided by CHFS are calculated to account for the sampling design, adjusting for unequal probabilities of selection at different stages.
 
-The household wealth variables consistently available in both 2017 and 2019 include total household assets and debts, financial assets and debts, agricultural assets and debts, business assets and debts, land, housing assets and debts, commercial property assets and debts, vehicle assets and debts, education debts, credit debts, medical debts, and other miscellaneous assets and debts.
-In 2019, an additional variable, garage assets, was reported separately from housing assets; this variable was not present in 2017.
+Several limitations of the CHFS dataset should be noted.
+First, it relies on self-reported values for income and assets, potentially introducing biases such as underreporting, recall inaccuracies, or privacy concerns [@gan_report_2014].
+Second, definitional changes across survey waves may affect comparability.
+For instance, the 2019 survey separately records garage assets, previously included within housing assets in the 2017 wave, potentially complicating longitudinal analyses.
+Additionally, certain variables in the dataset are top-coded to protect respondent privacy, with extreme values replaced by preset caps, accompanied by censoring indicators [@chfs_2019_2019].
+Such censoring is not accounted for in this study and might underestimate the level of income and wealth inequality in China.
 
-For consistency, this paper calculates equity values by deducting debts from corresponding assets, except for land, which is typically fully owned and non-transferable.
-Due to its relatively small aggregate value, garage assets in 2019 are recombined with housing assets for analytical consistency.
-
-Given the relatively short period between 2017 and 2019, this study aims to identify stable patterns in China's wealth inequality using data from these two years.
-Additionally, it compares the findings with earlier studies that utilize the same CHFS dataset to infer recent trends in wealth inequality[@tan_earnings_2017].
-
-# Analysis
-## Change of Gini from 2011 to 2019
-This section presents the Gini index analysis using the 2017 and 2019 CHFS data.
-The statistics from the 2011 CHFS data is from @tan_earnings_2017, which is work from CFHS's research team.
-All the income and wealth statistics are aggregated at the household level.
+The CHFS dataset has been extensively utilized in academic research.
+Studies leveraging the CHFS data have explored topics ranging from wealth distribution and inequality [@tan_earnings_2017], household finance in rural China [@gan_household_2014], household medical insurance expenditure [@li_household_2019], and many others.
 
@@
 
-{{< include summary_stats.txt >}}
+{{< include ../output/summary_stats.txt >}}
 
@@
 ## Factor decomposition of wealth inequality
-This section examines the composition and distribution of household wealth in China by analyzing specific asset and debt categories using data from the 2017 and 2019 China Household Finance Survey (CHFS).
-This detailed decomposition provides insights into the relative importance of different components contributing to wealth inequality.
+
+To understand the sources driving wealth inequality in China, this section applies the decomposition method developed by @lerman_income_1985.
+According to this method, the contribution of each wealth component to the overall Gini coefficient is decomposed as follows:
+
+$$
+G = \sum_{k=1}^{K} S_k G_k R_k
+$$
+
+where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth.
+This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking [@lerman_income_1985].
 
@@
 
-{{< include wealth_summary.txt >}}
+{{< include ../output/wealth_summary.txt >}}
 
@@
 
-The above table intuitively shows the state of household wealth allocation in China in 2017 and 2019.
-Clearly the mean and median value of house dominates other assets.
-It is also notable that despite the high inequality in housing assets, as it is shown in the next table, the proportion of households that have no houses at all is fairly low at just 10%.
-It is also surprising that the housing debt is very low compared to the asset value, and 82% of the households have no house debts.
+@tbl-wealth-summary summarizes household wealth allocation in China for 2017 and 2019.
+Housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types.
+Although housing significantly contributes to overall wealth inequality, homeownership remains widespread, with only about 10% of households reporting no housing assets.
+Moreover, housing-related debt remains minimal relative to housing values, with approximately 82% of households reporting no such debts, suggesting broad equity-based homeownership patterns.
 
-::: {#tbl-tbl2}
+Financial equity notably increased from 2017 to 2019, with mean values rising substantially from 103,355 RMB to 161,199 RMB and median values more than doubling from 17,170 RMB to 39,251 RMB.
+Correspondingly, financial assets increased their share in total household wealth from around 12% in 2017 to approximately 15% in 2019.
+Land asset ownership, reported by about 44% of households in 2017 and 41% in 2019, closely aligns with China's rural population proportion of approximately 42% in 2017 and 40% in 2019 [@world_bank_rural_nodate].
+Observed changes in land, commercial, and other assets should be interpreted cautiously, as it appears that some assets are redefined or newly added to the variables, which might partly reflect definitional adjustments rather than purely economic dynamics.
 
-{{< include 2017_gini_decom.txt >}}
+::: {#tbl-gini-decomposition-2017}
+
+{{< include ../output/2017_gini_decom.txt >}}
 
@@
 
-Decomposition of Wealth Gini Coefficient by Urban-Rural Division, 2017
+Factor Decomposition of Wealth Gini Coefficient, 2017
 :::
 
-::: {#tbl-tbl3}
+::: {#tbl-gini-decomposition-2019}
 
-{{< include 2019_gini_decom.txt >}}
+{{< include ../output/2019_gini_decom.txt >}}
 
@@
 
-Decomposition of Wealth Gini Coefficient by Urban-Rural Division, 2019
+Factor Decomposition of Wealth Gini Coefficient, 2019
 :::
 
-The factor decompostion of the Gini coefficient for household wealth in two years shows that housing asset dominates the wealth inequality in China.
-In both 2017 and 2019, the housing asset occupy over 70% of total household wealth and contribute to over 72% of the overall wealth Gini.
-This high contribution of housing equity to wealth inequality is consistent with other studies.
-@tan_earnings_2017 documented with the 2011 CHFS data that the housing asset to total asset ratio was 66.73%, and the 2017 and 2019 74.2% and 73% shows an sharp increase in the housing asset to total asset ratio from 2011 to 2019.
-Meanwhile, @kanbur_great_2021 used the 2002 and 2013 CHIP data and found that housing euqity contribues to 64% of wealth inequality in 2002 and 79% in 2013.
-Albeit variation in the statistics across sources, the results in @tbl-tbl3 demonstrates that in 2017 and 2019, the housing market is still the most important driver of wealth inequality in China.
-However, surprisingly,
+@tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019 show the factor decomposition results for 2017 and 2019, respectively.
+Both years confirm housing assets' persistent and dominant role in driving China's wealth inequality, contributing over 72% to the overall inequality.
+This aligns with earlier studies: @knight_why_2022 reported that housing equity's contribution rose significantly from 64% in 2002 to 79% in 2013, while @tan_earnings_2017 documented housing's share at approximately 67% of household wealth in 2011.
+Financial equity emerged as another important factor, increasing its contribution from approximately 11% in 2017 to about 14% in 2019.
+In contrast, contributions from commercial assets declined notably, potentially due to definitional changes in the survey.
+
+Overall, the structure of household wealth inequality in China remained relatively stable between 2017 and 2019.
+Housing assets continued to play a dominant role, with financial assets becoming increasingly significant, suggesting evolving household investment behaviors.
+The following part expands this analysis by conducting group-specific wealth inequality decompositions to understand the difference in wealth inequality across different rural and urban households and east and non-east regions.
+
+## Factor decomposition partitioned by rural and urban residence
+Since changes in the factor decomposition from 2017 to 2019 for rural and urban households closely mirror those observed for the entire population, only the 2019 decomposition results for rural and urban groups are presented here.
+
+::: {#tbl-urban-decom}
+
+{{< include ../output/2019_urban_decom.txt >}}
+
+Source: Author's calculation based on the 2019 CHFS data.
+
+Factor Decomposition of Wealth Gini Coefficient, 2019 (Urban)
+:::
+
+::: {#tbl-rural-decom}
+
+{{< include ../output/2019_rural_decom.txt >}}
+
+Source: Author's calculation based on the 2019 CHFS data.
+
+Factor Decomposition of Wealth Gini Coefficient, 2019 (Rural)
+:::
+
+The comparative factor decompositions for 2019 underscore several salient distinctions between the determinants of wealth inequality in urban and rural China.
+In both settings residential property remains the predominant source of inequality; nevertheless, its influence is markedly stronger in cities, where housing contributes roughly three‑quarters of the overall Gini coefficient, compared with just over one‑half in the countryside.
+This pronounced urban effect reflects the exceptionally high and spatially differentiated prices of metropolitan real estate as well as the tendency for affluent urban households to own multiple dwellings.
+
+By contrast, assets that are either absent from, or less significant in, urban portfolios assume a more prominent role in rural areas.
+Most notably, agricultural and homestead land accounts for almost one‑fifth of rural inequality but for only about six per cent of the urban total.
+The unequal distribution of land holdings, combined with large inter‑village variation in implicit land values, therefore constitutes a key driver of rural wealth dispersion.
+Commercial (business) assets exhibit a similar pattern: they explain roughly eleven per cent of rural inequality but just seven per cent in cities, mirroring the heterogeneity of small‑scale entrepreneurial activities in village and township economies.
+Conversely, financial equity is slightly more important in urban areas (around thirteen per cent) than in rural areas (about eleven per cent), consistent with better access of urban households to formal financial markets.
+
+Finally, debt liabilities play a modest equalising role in both residence types, as indicated by their negative contributions to the Gini.
+The mitigating effect is somewhat larger in urban China: mortgage debt alone reduces the urban coefficient by roughly 3.3 percentage points, relative to 2.1 percentage points in rural areas.
+Taken together, these findings imply that policy initiatives aimed at curbing wealth inequality ought to be tailored to local contexts.
+In cities, measures that temper speculative housing demand and widen access to diversified financial products are likely to prove most effective, whereas in rural regions reforms that strengthen land‑tenure security, facilitate land transactions, and support productive agricultural and non‑farm enterprises are likely to yield greater distributive gains.
+
+When benchmarked against the national decompositions for 2017 and 2019 (see \@tbl-gini-decomposition-2017 and \@tbl-gini-decomposition-2019), the urban–rural patterns largely reiterate the centrality of housing, land, and financial equity for Chinese wealth inequality.
+What the residence‑specific results add, however, is a clearer sense of _relative_ importance: housing is even more dominant in cities than the country‑wide average, whereas land and business assets loom larger in rural portfolios than the aggregate figures would suggest.
+At the same time, the contribution of financial assets in both residence types remains close to the national share, confirming that the rise of household financial investment is a broad‑based rather than location‑specific phenomenon.
+
+## Factor decomposition partitioned by east and non-east regions
+It is worth noting that from 2017 to 2019, housing's contribution to the Gini coefficient rose by 2.4 percentage points among households in the East region, whereas it decreased by 2.9 percentage points among Non-East households, despite housing's overall contribution remaining nearly identical across both years.
+This finding is significant, indicating that housing-related wealth inequality increased in East China but decreased in Non-East regions during this period.
+Aside from housing, contributions from other asset and debt categories showed similar trends for East and Non-East households.
+Therefore, only the decomposition results for 2019 are presented here.
+
+::: {#tbl-east-decom}
+
+{{< include ../output/2019_east_decom.txt >}}
+
+Source: Author's calculation based on the 2019 CHFS data.
+
+Factor Decomposition of Wealth Gini Coefficient, 2019 (East)
+:::
+
+::: {#tbl-noneast-decom}
+
+{{< include ../output/2019_noneast_decom.txt >}}
+
+Source: Author's calculation based on the 2019 CHFS data.
+
+Factor Decomposition of Wealth Gini Coefficient, 2019 (Non-East)
+:::
+
+Relative to the country‑wide decompositions for 2017 and 2019, the regional results confirm the pervasive importance of housing, land and financial assets in shaping wealth inequality across China.
+Yet they also reveal distinct regional configurations.
+In the coastal East, the contribution of housing climbs to more than 80 per cent of total inequality—roughly eight percentage points above the national average—underscoring the outsized role of metropolitan real‑estate markets.
+By contrast, in the Non‑East provinces housing accounts for just 58 per cent, while land and commercial (business) assets together explain nearly one‑quarter of inequality, double their combined share in the East.
+Financial equity is a sizeable driver in both regions, but its influence is marginally greater in the interior (≈ 16 %) than along the coast (≈ 12 %), reflecting the recent expansion of retail investment beyond the major coastal cities.
+Debt components continue to dampen inequality everywhere; however, the equalising effect of mortgage liabilities is slightly stronger in the East, where household borrowing is more closely tied to high‑value housing.
+These contrasts suggest that policies aimed at moderating wealth disparities will need to account for regional heterogeneity—addressing speculative housing dynamics in the East while simultaneously fostering land‑market reforms and business development in the interior—within the common framework of wider financial inclusion.
 
 ## Group decomposition of wealth inequality by rural and urban
-Group decomposition of wealth inequality by rural and urban
-
 To analyze the sources of wealth inequality, this paper employs the classic Gini decomposition identity initially developed by Bhattacharya and Mahalanobis (1967)[@bhattacharya_regional_1967].
@@
 
-{{< include rural_urban_summary.txt >}}
+{{< include ../output/rural_urban_summary.txt >}}
 
@@
 
-{{< include urban_rural_gini_decomp.txt >}}
+{{< include ../output/urban_rural_gini_decomp.txt >}}
 
@@
 
-{{< include region_summary.txt >}}
+{{< include ../output/region_summary.txt >}}
 
@@
 
-{{< include region_gini_decomp.txt >}}
+{{< include ../output/region_gini_decomp.txt >}}
 

### 2025-04-24 report/draft.qmd
update

@@
 # Introduction
-China has experienced remarkable economic growth in recent decades following its reform and opening-up policy initiated in 1978.
-However, this period has also been marked by a significant rise in income and wealth inequality [@piketty_capital_2019].
-Some scholars have provided evidence suggesting that China's overall inequality peaked around 2010 and subsequently began to decline [@kanbur_great_2021; @zhang_survey_2021].
-It is also well-documented that China's inequality is strongly intertwined with its urban-rural and coastal-inland divisions [@piketty_capital_2019; @zhang_survey_2021].
-This paper utilizes data from the 2017 and 2019 China Household Finance Survey (CHFS) to examine the current state of China's wealth inequality.
-Specifically, it investigates whether the latest publicly available data supports the view that China's inequality has continued to decline since around 2010, as well as how China's wealth inequality relates to urban-rural and regional disparities.
+China's remarkable economic growth over the past five decades was initiated by the reform and opening-up policy launched in 1978, transitioning the country from a centrally planned to a market-oriented economy.
+From the openning up in 1978 to 2024, China's GDP growth rate averaged over 9% per year, and 800 million people were lifted out of poverty, with considerable improvment in the nation's health and education [@world_bank_world_2024].
+This growth, however, has not been uniform across time, geography, or economic sectors.
+Coastal provinces such as Jiangsu and Guangdong experienced substantially higher growth rates, benefiting from early exposure to foreign investment and trade, while inland and rural areas lagged, leading to pronounced regional disparities [@liao_sixty_2016].
+Sectorally, China's strong growth has been driven by fixed-capital investment and export-oriented manufacuring, which althoug effective in the past, start to show signs of both domestic and international imbalances [@world_bank_world_2024].
+Since 2010, China's annual GDP growth rate has been declining from 10.6% in 2011 to 5.2% in 2023[@world_bank_gdp_2025].
+
+Concurrently, China experienced a substantial rise in wealth and income inequality.
+The wealth share controlled by the top 10% increased markedly from approximately 40% in the late 1970s to nearly 70% by 2015, while the share held by the bottom half significantly decreased [@piketty_capital_2019].
+Some studies indicate that China's overall inequality reached its peak around 2010, after which it started to gradually decline [@kanbur_great_2021; @zhang_survey_2021].
+Additionally, it is well-established that China's inequality closely correlates with its urban-rural and coastal-inland divisions, with urban and coastal regions exhibiting substantially higher levels of income and wealth compared to their rural and inland counterparts [@piketty_capital_2019; @zhang_survey_2021].
+
+This paper analyzes the contemporary landscape of wealth inequality in China using recent data from the 2017 and 2019 China Household Finance Survey (CHFS).
+It specifically investigates whether recent trends confirm claims of continued inequality reduction since 2010.
+Through factor decomposition, the study examines the sources of wealth inequality within the entire population as well as separately within urban-rural and coastal-inland subgroups, highlighting variations in inequality among these different groups.
+Furthermore, the study applies group decomposition methods to assess the extent to which rural-urban and coastal-inland disparities contribute to overall wealth inequality.
+Through these detailed decompositions and analyses, the paper aims to provide a comprehensive and nuanced understanding of current patterns and dynamics in China's wealth inequality.
 
@@
 They find the national Gini coefficient peaked at about 0.525 in 2010, subsequently declining to approximately 0.476 in 2016 before rising slightly again.
-Consistent with Zhang (2021), they attribute this recent turnaround primarily to narrowing rural-urban and coastal-inland disparities, driven by demographic changes, tightening rural labor markets, and rising rural wages.
+Consistent with @zhang_survey_2021, they attribute this recent turnaround primarily to narrowing rural-urban and coastal-inland disparities, driven by demographic changes, tightening rural labor markets, and rising rural wages.
 
@@
 The CHFS dataset has been extensively utilized in academic research.
-Studies leveraging the CHFS data have explored topics ranging from wealth distribution and inequality [@tan_earnings_2017], household finance in rural China [@gan_household_2014], household medical insurance expenditure [@li_household_2019], and many others.
+Studies leveraging the CHFS data have explored topics ranging from wealth distribution and inequality [@tan_earnings_2017], household finance in rural China [@gan_report_2014], household medical insurance expenditure [@liu_health_2022], and many others.
 
+# Analysis
+## Summary Statistics
 ::: {#tbl-tbl1}
@@
 
-Notably, the growth rate of mean and median income from 2011 to 2017 is clearly higher than that of wealth but from 2017 to 2019, the growth of mean and median wealth surpasses that of income by a great margin.
+Notably, while the growth rate of mean and median income from 2011 to 2017 substantially outpaced that of wealth, this pattern reversed between 2017 and 2019, where the growth of mean and median wealth greatly surpassed income growth.
+This later-period divergence may reflect factors such as significant asset price appreciation relative to income gains.
 
@@
 Land asset ownership, reported by about 44% of households in 2017 and 41% in 2019, closely aligns with China's rural population proportion of approximately 42% in 2017 and 40% in 2019 [@world_bank_rural_nodate].
+However, later results in this paper show that the gini coefficient for land in the rural area was 89.7% in the rural area in 2019, which is significantly higher than the gini coefficient for wealth and for housing in rural areas.
+This shows that dispite a wide distribution of land ownership, the distribution of land is highly unequal in modern China.
+
 Observed changes in land, commercial, and other assets should be interpreted cautiously, as it appears that some assets are redefined or newly added to the variables, which might partly reflect definitional adjustments rather than purely economic dynamics.
@@
 @tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019 show the factor decomposition results for 2017 and 2019, respectively.
-Both years confirm housing assets' persistent and dominant role in driving China's wealth inequality, contributing over 72% to the overall inequality.
+Both years confirm housing assets' persistent and dominant role in driving China's wealth inequality, contributing over 72% to the overall inequality (72.7% in 2017, 72.3% in 2019).
 This aligns with earlier studies: @knight_why_2022 reported that housing equity's contribution rose significantly from 64% in 2002 to 79% in 2013, while @tan_earnings_2017 documented housing's share at approximately 67% of household wealth in 2011.
-Financial equity emerged as another important factor, increasing its contribution from approximately 11% in 2017 to about 14% in 2019.
+Financial equity emerged as another important factor, increasing its contribution from approximately 11.5% in 2017 to about 13.5% in 2019.
 In contrast, contributions from commercial assets declined notably, potentially due to definitional changes in the survey.
@@
 Housing assets continued to play a dominant role, with financial assets becoming increasingly significant, suggesting evolving household investment behaviors.
-The following part expands this analysis by conducting group-specific wealth inequality decompositions to understand the difference in wealth inequality across different rural and urban households and east and non-east regions.
+To explore heterogeneity in these patterns, the following sections decompose wealth inequality within specific demographic subgroups.
 
@@
 
-When benchmarked against the national decompositions for 2017 and 2019 (see \@tbl-gini-decomposition-2017 and \@tbl-gini-decomposition-2019), the urban–rural patterns largely reiterate the centrality of housing, land, and financial equity for Chinese wealth inequality.
+When benchmarked against the national decompositions for 2017 and 2019 (see @tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019), the urban–rural patterns largely reiterate the centrality of housing, land, and financial equity for Chinese wealth inequality.
 What the residence‑specific results add, however, is a clearer sense of _relative_ importance: housing is even more dominant in cities than the country‑wide average, whereas land and business assets loom larger in rural portfolios than the aggregate figures would suggest.
@@
 
-Relative to the country‑wide decompositions for 2017 and 2019, the regional results confirm the pervasive importance of housing, land and financial assets in shaping wealth inequality across China.
-Yet they also reveal distinct regional configurations.
-In the coastal East, the contribution of housing climbs to more than 80 per cent of total inequality—roughly eight percentage points above the national average—underscoring the outsized role of metropolitan real‑estate markets.
-By contrast, in the Non‑East provinces housing accounts for just 58 per cent, while land and commercial (business) assets together explain nearly one‑quarter of inequality, double their combined share in the East.
-Financial equity is a sizeable driver in both regions, but its influence is marginally greater in the interior (≈ 16 %) than along the coast (≈ 12 %), reflecting the recent expansion of retail investment beyond the major coastal cities.
-Debt components continue to dampen inequality everywhere; however, the equalising effect of mortgage liabilities is slightly stronger in the East, where household borrowing is more closely tied to high‑value housing.
-These contrasts suggest that policies aimed at moderating wealth disparities will need to account for regional heterogeneity—addressing speculative housing dynamics in the East while simultaneously fostering land‑market reforms and business development in the interior—within the common framework of wider financial inclusion.
+Relative to the country‑wide decompositions for 2017 and 2019, the regional results confirm the pervasive importance of housing, land and financial assets in shaping wealth inequality across China, yet they also reveal distinct regional configurations.
+In the coastal East, housing's influence is particularly pronounced; its contribution climbs to over 80 per cent (80.3%) of total inequality—roughly eight percentage points above the national average—underscoring the outsized role of metropolitan real‑estate markets and its strong alignment with overall wealth rank (Gini correlation 0.96).
+By contrast, in the Non‑East provinces housing accounts for a lower share at 58.3 per cent.
+Here, land and commercial (business) assets play a much larger role, together explaining nearly one-quarter of inequality (23.2%, combining 13.8% from land and 9.4% from commercial), substantially higher than their combined share of about 7.6% in the East.
+This reflects not only portfolio differences but also stronger correlations between these assets and wealth rank in the Non-East (Gini correlations of 0.83 for land and 0.80 for commercial, vs.
+0.35 and 0.72 respectively in the East).
+
+Financial equity is a sizeable driver in both regions, though its contribution and correlation with wealth rank are marginally greater in the interior (15.7% contribution, 0.78 correlation) than along the coast (12.4% contribution, 0.77 correlation), perhaps reflecting the recent expansion of retail investment beyond major coastal cities.
+Debt components, particularly mortgage liabilities, serve to dampen inequality in both regions through their negative contributions to the overall Gini coefficient.
+These contrasts underscore that policies aimed at moderating wealth disparities need to account for regional heterogeneity—potentially addressing speculative housing dynamics in the East while simultaneously fostering land‑market reforms and business development in the interior—within a common framework of wider financial inclusion.
 
@@
 The summary statistics highlight significant and persistent disparities between urban and rural China.
-Urban households consistently exhibit higher mean and median incomes and wealth compared to rural households.
+Urban households consistently exhibit higher mean incomes and wealth compared to rural households.
 Urban mean wealth was 3.48 times greater than rural mean wealth in 2017 and widened slightly to 3.78 times in 2019.
-Median urban wealth was 2.82 times higher than the rural median in 2017, though this ratio decreased slightly to 2.52 in 2019.
-Additionally, the Gini coefficient for wealth is consistently higher in rural areas, though both rural and urban Gini coefficients declined modestly between 2017 and 2019.
+Additionally, the Gini coefficient for wealth is consistently higher in rural areas (0.6945 in 2017, 0.6629 in 2019) compared to urban areas (0.6573 in 2017, 0.6165 in 2019), though Gini coefficients for both groups declined modestly between the two years.
 
@@
 
-The decomposition results indicate that about 55–57% of total wealth inequality is attributable to within-group disparities among urban and rural households.
-The between-group component, reflecting the urban-rural wealth divide, contributes approximately 33–34%.
-The overlap component, representing inequality due to intersecting distributions of urban and rural wealth, accounts for around 9–12% of total inequality.
+The decomposition results indicate that about 55–57% (55.4% in 2017, 56.8% in 2019) of total wealth inequality is attributable to within-group disparities among urban and rural households.
+The between-group component, reflecting the urban-rural wealth divide, contributes approximately 33–34% (32.6% in 2017, 33.7% in 2019).
+The overlap component, representing inequality due to intersecting distributions of urban and rural wealth, accounts for around 9–12% (12.0% in 2017, 9.5% in 2019) of total inequality.
 The consistency of these shares between 2017 and 2019 underscores the continued significance of the urban-rural divide in shaping China's wealth inequality.
@@
 The above summary statistics shows that the mean and median wealth in the East is significantly higher than in the other regions in both 2017 and 2019.
-Notably, the mean and median wealth of the West-area households surpassed that of the Middle in 2019, but West's Gini coefficient rose to the highest among all regions in 2019 as well.
+Notably, the mean wealth of West-area households surpassed that of the Middle region in 2019, although its median wealth remained lower.
+Concurrently, the West's Gini coefficient rose to become the highest among all regions in 2019 (0.6947).
 
@@
 
-The group decompsition shows that similar to the urban-rural decomposition, the between-group component contribues over 30% to the overall wealth inequality in both 2017 and 2019.
-The between group contribution saw a slight decrease from 2017 to 2019 with a similari increase in the residual component.
+The group decomposition shows that, similar to the urban-rural decomposition, the between-group component contributes over 30% to overall wealth inequality in both 2017 and 2019 (36.9% and 33.8% respectively).
+The between-group contribution saw a slight decrease from 2017 to 2019, with a similar increase in the residual component (from 32.0% to 34.5%).
 This might signal some modest convergence in household wealth between regions.
 
+## Conclusion
+This paper analyzed wealth inequality in China using the 2017 and 2019 CHFS data, finding a modest decline in the overall Gini coefficient (0.701 to 0.667), consistent with literature suggesting a post-2010 peak [@kanbur_great_2021; @zhang_survey_2021].
+Factor decomposition confirmed housing assets as the primary driver (contributing over 72%), aligning with previous studies [@knight_why_2022; @tan_earnings_2017], while financial equity's contribution grew (from approximately 11% to 14%).
+Partitioned analysis highlighted housing's stronger impact in urban areas (around 75% vs.
+approximately 52% rural) and especially the East region (over 80% vs.
+approximately 58% Non-East), with divergent trends emerging between 2017-2019 (increasing contribution in the East, decreasing elsewhere).
+Group decomposition further revealed that while within-group inequality dominates (accounting for approximately 55-60%), substantial inequality stems from persistent urban-rural (around 33%) and regional (over 30%) divides.
+
+These dynamics likely reflect intertwined factors: rapid urbanization, uneven real estate appreciation and speculation impacting housing's role; financial market development driving the rise of financial assets, potentially favoring wealthier urban households; and deep-rooted structural issues like historical policies, the Hukou system, and differential access to opportunities maintaining the significant urban-rural and regional gaps, despite potential slow convergence hinted by recent trends and policies [@kanbur_great_2021].
+Consequently, addressing China's high wealth inequality requires a nuanced approach beyond national measures, necessitating context-specific policies such as targeted housing market regulation in the East, support for land and business development in rural and Non-East areas, and broader efforts towards financial inclusion across the country.
+
 # References

### 2025-04-25 report/draft.qmd
update

@@
 # Introduction
-China's remarkable economic growth over the past five decades was initiated by the reform and opening-up policy launched in 1978, transitioning the country from a centrally planned to a market-oriented economy.
-From the openning up in 1978 to 2024, China's GDP growth rate averaged over 9% per year, and 800 million people were lifted out of poverty, with considerable improvment in the nation's health and education [@world_bank_world_2024].
-This growth, however, has not been uniform across time, geography, or economic sectors.
-Coastal provinces such as Jiangsu and Guangdong experienced substantially higher growth rates, benefiting from early exposure to foreign investment and trade, while inland and rural areas lagged, leading to pronounced regional disparities [@liao_sixty_2016].
-Sectorally, China's strong growth has been driven by fixed-capital investment and export-oriented manufacuring, which althoug effective in the past, start to show signs of both domestic and international imbalances [@world_bank_world_2024].
-Since 2010, China's annual GDP growth rate has been declining from 10.6% in 2011 to 5.2% in 2023[@world_bank_gdp_2025].
 
-Concurrently, China experienced a substantial rise in wealth and income inequality.
-The wealth share controlled by the top 10% increased markedly from approximately 40% in the late 1970s to nearly 70% by 2015, while the share held by the bottom half significantly decreased [@piketty_capital_2019].
-Some studies indicate that China's overall inequality reached its peak around 2010, after which it started to gradually decline [@kanbur_great_2021; @zhang_survey_2021].
-Additionally, it is well-established that China's inequality closely correlates with its urban-rural and coastal-inland divisions, with urban and coastal regions exhibiting substantially higher levels of income and wealth compared to their rural and inland counterparts [@piketty_capital_2019; @zhang_survey_2021].
+China's economic trajectory over the past five decades, initiated by the reform and opening-up policy launched in 1978, involved a transition from a centrally planned to a more market-oriented economy.
+This period yielded high average GDP growth rates alongside considerable improvements in poverty reduction, health, and education [@world_bank_world_2024; @ravallion_chinas_2009].
+This growth, however, manifested unevenly across time, geography, and economic sectors.
+Coastal provinces, for instance, often experienced higher growth rates, benefiting from earlier integration with global markets, while inland and rural areas developed more slowly, contributing to regional disparities [@liao_sixty_2016].
+Sectorally, growth was initially propelled by fixed-capital investment and export-oriented manufacturing, a model that later showed signs of creating imbalances [@world_bank_world_2024].
+China's annual GDP growth rate moderated from 10.6% in 2010 to 5.2% by 2023, reflecting a slowdown from the high rates experienced earlier in the decade [@world_bank_gdp_2025].
 
-This paper analyzes the contemporary landscape of wealth inequality in China using recent data from the 2017 and 2019 China Household Finance Survey (CHFS).
-It specifically investigates whether recent trends confirm claims of continued inequality reduction since 2010.
-Through factor decomposition, the study examines the sources of wealth inequality within the entire population as well as separately within urban-rural and coastal-inland subgroups, highlighting variations in inequality among these different groups.
-Furthermore, the study applies group decomposition methods to assess the extent to which rural-urban and coastal-inland disparities contribute to overall wealth inequality.
-Through these detailed decompositions and analyses, the paper aims to provide a comprehensive and nuanced understanding of current patterns and dynamics in China's wealth inequality.
+Parallel to this economic expansion, China experienced a substantial increase in wealth and income inequality.
+Analysis combining diverse data sources reveals a marked shift in distribution between 1978 and 2015; the wealth share controlled by the top 10% increased from approximately 40% to nearly 70%, while the share held by the bottom 50% diminished considerably [@piketty_capital_2019].
+Wealth inequality escalated rapidly, with the Gini coefficient for household wealth per capita rising markedly between 2002 and 2013, reaching high levels by international comparison [@knight_why_2022; @tan_earnings_2017].
+Some research suggests that overall income inequality may have reached a peak around 2008-2010, potentially followed by a gradual decline or stabilization, linked partly to narrowing structural divides [@zhang_survey_2021; @kanbur_great_2021].
+It is well-documented that China's inequality dynamics intersect closely with urban-rural and coastal-inland divisions; urban and coastal regions consistently exhibit higher average levels of income and wealth compared to their rural and inland counterparts [@piketty_capital_2019; @zhang_survey_2021].
+The urban-rural income ratio remained high even in the late 2010s, despite some narrowing from its 2009 peak, and regional income differences persist [@zhang_survey_2021], findings consistent across multiple data sources [@xie_income_2014].
 
-# Background
-This section presents some notable findings on China's income and wealth inequality in the existing literature.
+This study contributes to the understanding of China's wealth inequality by employing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS).
+The analysis focuses specifically on assessing whether wealth inequality trends during this later period substantiate claims of ongoing inequality reduction observed in earlier research focused primarily on income.
+The paper utilizes factor decomposition methods to pinpoint the contributions of various asset types to overall wealth inequality.
+A further level of analysis involves applying this factor decomposition separately to distinct population subgroups defined by urban-rural residence and coastal-inland location, allowing for an examination of potential heterogeneity in the drivers of inequality across these dimensions.
+Additionally, group decomposition techniques are used to quantify the portion of total wealth inequality attributable to the average wealth differences between these urban-rural and coastal-inland groups.
+This combination of recent data and multiple decomposition strategies aims to offer a detailed perspective on the patterns and structural underpinnings of wealth inequality in contemporary China.
 
-China has experienced a considerable increase in income and wealth inequality since the reform and opening-up policy initiated in 1978.
-Using diverse sources, including tax records, household surveys, and private wealth rankings, @piketty_capital_2019 provide a detailed analysis of the transformation in China's inequality profile.
-Their findings show that between 1978 and 2015, the income share of the top 10% increased from 27% to 41%, while the share of the bottom 50% decreased from 27% to 15%.
-Wealth concentration also intensified significantly: the wealth share held by the top 10% rose from around 40% to nearly 70%, whereas the bottom 50%'s share dropped from over 15% to about 5%.
-Additionally, they documented an increase in the urban-to-rural per capita income ratio, from less than 200% in 1978 to approximately 350% by 2015.
-Humorously noting this stark shift, they remarked that China had transitioned from Nordic-level equality to U.S.-level inequality during the reform era.
+# Literature Review
+This section reviews notable findings from existing literature on the evolution of China's income and wealth inequality following the market-oriented reforms initiated in 1978.
 
-@knight_why_2022 further document a rapid rise in China's household wealth inequality between 2002 and 2013, reporting that the national Gini coefficient for household wealth per capita increased markedly from 0.50 to 0.62.
-They identify housing wealth as the primary driver behind this increase, with its share rising from 53% to 73% of total household wealth, and its contribution to overall wealth inequality growing from 64% to 79%.
-The authors emphasize that differential savings rates across income groups, uneven house price inflation—particularly pronounced in major urban centers—and widening urban-rural disparities significantly intensified wealth inequality during this period.
+China's economic trajectory since 1978 has been characterized by high rates of growth alongside a considerable increase in economic disparities.
+Utilizing diverse data including tax records, surveys, and wealth rankings, @piketty_capital_2019 meticulously document this transformation.
+Their analysis reveals a substantial redistribution of national income between 1978 and 2015: the top 10% saw their income share rise from 27% to 41%, while the bottom 50%'s share conversely fell from 27% to 15%.
+Wealth concentration followed a similar, potentially more pronounced, pattern.
+The wealth share of the top 10% increased from approximately 40% to nearly 70%, diminishing the bottom 50%'s share from over 15% to around 5%.
+Concurrently, the urban-rural income gap widened, with the average per capita income ratio increasing from less than 2:1 to about 3.5:1.
+This shift led @piketty_capital_2019 to observe that China moved from equality levels comparable to Nordic countries towards those observed in the U.S.
+during this period.
+Complementing this, @xie_income_2014 established, using multiple surveys from the mid-2000s onwards, that China's Gini coefficient had reached levels between 0.53 and 0.55, exceeding official figures and the US benchmark at the time.
 
-The income and wealth inequality highly intersects with urban-rural and reginal disparities.
-@zhang_survey_2021 highlights significant rural-urban and regional disparities underlying China's rising inequality since its economic reforms.
-The urban-rural income ratio increased sharply, peaking at around 3.3 in 2009 before declining slightly to 2.7 by 2019, still substantially higher than pre-reform levels.
-Despite a notable decline in overall inequality after 2010—with the national Gini coefficient decreasing from a peak of 0.491 in 2008 to approximately 0.465 in 2019—regional inequality remains pronounced.
-For instance, per capita GDP in the eastern coastal region was roughly 2.6 times higher than in the western region by 2018.
-Additionally, rural-to-urban migration grew dramatically from fewer than 20 million in 1990 to approximately 290 million by 2019, reflecting both the economic disparity between regions and the persistent urban-rural divide, exacerbated by restricted access to urban public services under the Hukou household registration system.
+Focusing on the period between 2002 and 2013, @knight_why_2022 provide further evidence on the escalation of household wealth inequality, reporting an increase in the Gini coefficient for wealth per capita from 0.50 to 0.62.
+A key finding is the role of housing wealth, which grew from 53% to 73% of total household assets and accounted for an increasing share of wealth inequality, rising from 64% to 79% of the total.
+The authors attribute the intensification of wealth inequality during these years to a combination of factors, including differential saving rates favoring the wealthy, and uneven house price inflation which benefited homeowners in major urban centers.
 
-Consistent with the finding in @zhang_survey_2021 that China's inequality peaked in 2008 and 2009, @kanbur_great_2021 similarly document that China's inequality increased sharply after economic reforms but started to plateau around 2010.
-They find the national Gini coefficient peaked at about 0.525 in 2010, subsequently declining to approximately 0.476 in 2016 before rising slightly again.
-Consistent with @zhang_survey_2021, they attribute this recent turnaround primarily to narrowing rural-urban and coastal-inland disparities, driven by demographic changes, tightening rural labor markets, and rising rural wages.
+These trends in income and wealth inequality are deeply intertwined with persistent structural divides within China.
+Significant rural-urban and regional disparities remain central to understanding China's inequality landscape [@zhang_survey_2021; @xie_income_2014].
+The urban-rural income ratio, despite some decline after peaking around 3.3 in 2009, remained elevated at approximately 2.7 in 2019 [@zhang_survey_2021].
+This gap is reinforced by the *hukou* (household registration) system, which limits rural migrants' access to urban services and opportunities, even as migration increased substantially from 1990 to 2019.
+Regional inequality also persists, with coastal areas generally outpacing inland regions [@liao_sixty_2016].
+While overall income inequality showed signs of stabilization or decline after reaching a peak around 2008-2010 [@zhang_survey_2021; @kanbur_great_2021], attributed partly to narrowing structural gaps driven by demographic shifts and tightening rural labor markets, geographical disparities endure.
+
+The transition towards a market economy itself introduced multiple drivers of inequality [@benjamin_income_2008; @zhang_survey_2021].
+In rural China, the rise was associated with unequal access to non-agricultural income sources, particularly family businesses, alongside periods of slow growth or falling prices in the agricultural sector [@benjamin_income_2008; @ravallion_chinas_2009].
+In urban areas, contributing factors included the phasing out of subsidies and entitlements, increasing wage dispersion driven by market forces and enterprise reforms, higher returns to education (especially tertiary degrees), and unemployment resulting from State-Owned Enterprise (SOE) restructuring [@benjamin_income_2008].
+Uneven regional development was further amplified by policies favoring coastal regions, as well as the differential impacts of globalization and marketization [@liao_sixty_2016].
+It is worth noting, however, that considerable inequality also exists within specific localities, suggesting that factors beyond broad regional or urban-rural classifications are also influential [@benjamin_income_2008].
+
+Furthermore, the pattern of economic growth influenced distributional outcomes.
+Research indicates that growth originating in the agricultural and rural sectors was more effective at reducing both poverty and overall inequality compared to growth in the industrial or service sectors [@ravallion_chinas_2009].
+This suggests the importance of the initial economic structure and the specific mechanisms linking growth to income gains across population segments.
+Conversely, the rise in inequality itself appeared to impede poverty reduction; provinces that began the reform period with higher levels of inequality experienced slower subsequent progress in reducing poverty, partly because growth was less effective at lifting incomes at the lower end of the distribution in these areas [@ravallion_chinas_2009].
+Specific policies, such as the initial agrarian reforms and adjustments to agricultural procurement prices, also demonstrably affected poverty and inequality levels.
+While income inequality may have plateaued or slightly decreased post-2010, linked to narrowing structural gaps and potentially policy interventions [@zhang_survey_2021; @liao_sixty_2016], wealth concentration, influenced by real estate appreciation, continued its upward trend well into the 2010s [@knight_why_2022; @piketty_capital_2019].
 
 # Data
-The analysis in this study utilizes data from the 2017 and 2019 waves of the China Household Finance Survey (CHFS), a nationally representative dataset collected by the Survey and Research Center for China Household Finance at Southwestern University of Finance and Economics [@gan_report_2014].
-The CHFS dataset includes household-level financial information on income, assets, debts, consumption, and demographic characteristics.
-Data are collected through face-to-face interviews using a multi-stage probability proportional to size (PPS) sampling method, ensuring representativeness at both national and provincial levels.
-Sample weights provided by CHFS are calculated to account for the sampling design, adjusting for unequal probabilities of selection at different stages.
+The empirical analysis in this paper utilizes microdata from the 2017 and 2019 waves of the China Household Finance Survey (CHFS).
+The CHFS, conducted biennially by the Survey and Research Center for China Household Finance at Southwestern University of Finance and Economics (SWUFE) since 2011, is designed to collect household financial information representative at both national and provincial levels [@gan_report_2014].
+The survey employs a stratified, three-stage Probability Proportional to Size (PPS) random sampling methodology.
+This approach involves selecting counties, followed by village or neighborhood committees, and then households, incorporating oversampling of high-wealth households to better capture asset distribution.
 
-Several limitations of the CHFS dataset should be noted.
-First, it relies on self-reported values for income and assets, potentially introducing biases such as underreporting, recall inaccuracies, or privacy concerns [@gan_report_2014].
-Second, definitional changes across survey waves may affect comparability.
-For instance, the 2019 survey separately records garage assets, previously included within housing assets in the 2017 wave, potentially complicating longitudinal analyses.
-Additionally, certain variables in the dataset are top-coded to protect respondent privacy, with extreme values replaced by preset caps, accompanied by censoring indicators [@chfs_2019_2019].
-Such censoring is not accounted for in this study and might underestimate the level of income and wealth inequality in China.
+Data are gathered through face-to-face interviews using Computer-Assisted Personal Interviewing (CAPI), collecting information at both household and individual levels.
+As with most survey data, CHFS figures are subject to potential limitations.
+Asset and income data rely on self-reporting, which can be affected by reporting error or bias.
+Respondent privacy is protected through methods including the top-coding of certain high-value income and asset variables; this may result in an underestimation of the upper tails of the respective distributions and potentially impact inequality metrics.
+Researchers should also note that minor adjustments in variable definitions between the 2017 and 2019 waves, such as the separate classification of garage assets in 2019, may require consideration for longitudinal comparisons.
+Missing data are handled through specific coding and imputation for certain variables.
 
-The CHFS dataset has been extensively utilized in academic research.
-Studies leveraging the CHFS data have explored topics ranging from wealth distribution and inequality [@tan_earnings_2017], household finance in rural China [@gan_report_2014], household medical insurance expenditure [@liu_health_2022], and many others.
+The CHFS provides sample weights, calculated based on the multi-stage sampling design and adjusted using population benchmarks, to facilitate representative analysis.
+The dataset has been utilized in numerous academic studies exploring various facets of the Chinese economy and household behavior.
+Prior research using CHFS data has examined topics such as income and wealth distribution [@tan_earnings_2017], rural household finance [@gan_report_2014], household responses to health shocks and insurance availability [@liu_health_2022] and many others [@tan_earnings_2017; @gan_report_2014; @liu_health_2022].
+
+The 2017 CHFS wave surveyed approximately 38,994 households, while the 2019 wave included 34,643 households across 343 county-level units in 29 provinces [@chfs_2017_2017; @chfs_2019_2019].
+For the present analysis, observations identified as low quality in the 2017 data, along with those lacking essential variables for this study, were omitted; however, the final analytical sample retains the substantial majority of the original observations.
 

### 2025-04-26 report/draft.qmd
update

@@
 China's economic trajectory over the past five decades, initiated by the reform and opening-up policy launched in 1978, involved a transition from a centrally planned to a more market-oriented economy.
-This period yielded high average GDP growth rates alongside considerable improvements in poverty reduction, health, and education [@world_bank_world_2024; @ravallion_chinas_2009].
+This period yielded high average GDP growth rates averaged over 9% per year along with considerable improvements in poverty reduction, health, and education [@world_bank_world_2024].
 This growth, however, manifested unevenly across time, geography, and economic sectors.
@@
 
-Parallel to this economic expansion, China experienced a substantial increase in wealth and income inequality.
-Analysis combining diverse data sources reveals a marked shift in distribution between 1978 and 2015; the wealth share controlled by the top 10% increased from approximately 40% to nearly 70%, while the share held by the bottom 50% diminished considerably [@piketty_capital_2019].
-Wealth inequality escalated rapidly, with the Gini coefficient for household wealth per capita rising markedly between 2002 and 2013, reaching high levels by international comparison [@knight_why_2022; @tan_earnings_2017].
-Some research suggests that overall income inequality may have reached a peak around 2008-2010, potentially followed by a gradual decline or stabilization, linked partly to narrowing structural divides [@zhang_survey_2021; @kanbur_great_2021].
-It is well-documented that China's inequality dynamics intersect closely with urban-rural and coastal-inland divisions; urban and coastal regions consistently exhibit higher average levels of income and wealth compared to their rural and inland counterparts [@piketty_capital_2019; @zhang_survey_2021].
-The urban-rural income ratio remained high even in the late 2010s, despite some narrowing from its 2009 peak, and regional income differences persist [@zhang_survey_2021], findings consistent across multiple data sources [@xie_income_2014].
+Parallel to this economic expansion, China experienced a significant rise in wealth and income inequality.
+@piketty_capital_2019 reveals a marked shift in wealth distribution between 1978 and 2015, highlighting that the wealth share controlled by the top 10% increased substantially, from approximately 40% to nearly 70%, while the share held by the bottom 50% diminished significantly.
+This trend of rising inequality is further corroborated by the Gini coefficient for household wealth per capita, which increased notably from 0.5 in 2002 to 0.63 in 2013, according to an official national survey [@knight_why_2022].
+Such levels of inequality are notably high by international standards [@tan_earnings_2017].
+Furthermore, China's inequality dynamics are closely intertwined with urban-rural and coastal-inland divisions; urban and coastal regions consistently exhibit higher average income and wealth levels compared to their rural and inland counterparts [@piketty_capital_2019; @zhang_survey_2021].
+Despite some narrowing since its peak in 2009, the urban-rural income ratio remained high in the late 2010s, and significant regional income disparities persist [@zhang_survey_2021], findings consistent across multiple data sources [@xie_income_2014].
 
 This study contributes to the understanding of China's wealth inequality by employing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS).
-The analysis focuses specifically on assessing whether wealth inequality trends during this later period substantiate claims of ongoing inequality reduction observed in earlier research focused primarily on income.
+The analysis focuses specifically on assessing whether wealth inequality trends during this later period substantiate claims of ongoing inequality reduction observed in earlier research focused primarily on income [@kanbur_great_2021; @zhang_survey_2021].
 The paper utilizes factor decomposition methods to pinpoint the contributions of various asset types to overall wealth inequality.
@@
 
-Focusing on the period between 2002 and 2013, @knight_why_2022 provide further evidence on the escalation of household wealth inequality, reporting an increase in the Gini coefficient for wealth per capita from 0.50 to 0.62.
-A key finding is the role of housing wealth, which grew from 53% to 73% of total household assets and accounted for an increasing share of wealth inequality, rising from 64% to 79% of the total.
-The authors attribute the intensification of wealth inequality during these years to a combination of factors, including differential saving rates favoring the wealthy, and uneven house price inflation which benefited homeowners in major urban centers.
+Focusing on the period between 2002 and 2013, @knight_why_2022 provide further evidence on the escalation of household wealth inequality using the Chinese Household Income Project (CHIP) data, reporting an increase in the Gini coefficient for wealth per capita from 0.50 to 0.62.
+A significant contributor to this increase was housing wealth, which expanded from 53% to 73% of total household assets.
+Housing also became increasingly central to wealth inequality, with its contribution rising from 64% to 79% of the overall inequality.
+The authors attribute the intensification of wealth inequality during these years to factors such as differential saving rates favoring the wealthy and uneven house price inflation that particularly benefited homeowners in major urban centers.
 
@@
 Regional inequality also persists, with coastal areas generally outpacing inland regions [@liao_sixty_2016].
-While overall income inequality showed signs of stabilization or decline after reaching a peak around 2008-2010 [@zhang_survey_2021; @kanbur_great_2021], attributed partly to narrowing structural gaps driven by demographic shifts and tightening rural labor markets, geographical disparities endure.
+While overall income inequality showed signs of stabilization or decline after reaching a peak around 2008-2010 [@kanbur_great_2021; @zhang_survey_2021], attributed partly to narrowing structural gaps driven by demographic shifts and tightening rural labor markets, geographical disparities endure.
 
@@
 
-Furthermore, the pattern of economic growth influenced distributional outcomes.
-Research indicates that growth originating in the agricultural and rural sectors was more effective at reducing both poverty and overall inequality compared to growth in the industrial or service sectors [@ravallion_chinas_2009].
-This suggests the importance of the initial economic structure and the specific mechanisms linking growth to income gains across population segments.
-Conversely, the rise in inequality itself appeared to impede poverty reduction; provinces that began the reform period with higher levels of inequality experienced slower subsequent progress in reducing poverty, partly because growth was less effective at lifting incomes at the lower end of the distribution in these areas [@ravallion_chinas_2009].
-Specific policies, such as the initial agrarian reforms and adjustments to agricultural procurement prices, also demonstrably affected poverty and inequality levels.
-While income inequality may have plateaued or slightly decreased post-2010, linked to narrowing structural gaps and potentially policy interventions [@zhang_survey_2021; @liao_sixty_2016], wealth concentration, influenced by real estate appreciation, continued its upward trend well into the 2010s [@knight_why_2022; @piketty_capital_2019].
-
 # Data
@@
 
-Data are gathered through face-to-face interviews using Computer-Assisted Personal Interviewing (CAPI), collecting information at both household and individual levels.
+Data are gathered through face-to-face interviews using Computer-Assisted Personal Interviewing (CAPI), collecting information at both household and individual levels [@gan_report_2014].
 As with most survey data, CHFS figures are subject to potential limitations.
@@
 Respondent privacy is protected through methods including the top-coding of certain high-value income and asset variables; this may result in an underestimation of the upper tails of the respective distributions and potentially impact inequality metrics.
-Researchers should also note that minor adjustments in variable definitions between the 2017 and 2019 waves, such as the separate classification of garage assets in 2019, may require consideration for longitudinal comparisons.
-Missing data are handled through specific coding and imputation for certain variables.
+It is also worth noting that potential adjustments in variable definitions between the 2017 and 2019 waves, such as the separate classification of garage assets in 2019 which was not present in 2017, may require consideration for longitudinal comparisons.
 
@@
 The dataset has been utilized in numerous academic studies exploring various facets of the Chinese economy and household behavior.
-Prior research using CHFS data has examined topics such as income and wealth distribution [@tan_earnings_2017], rural household finance [@gan_report_2014], household responses to health shocks and insurance availability [@liu_health_2022] and many others [@tan_earnings_2017; @gan_report_2014; @liu_health_2022].
+Prior research using CHFS data has examined topics such as income and wealth distribution [@tan_earnings_2017], rural household finance [@gan_report_2014], household responses to health shocks and insurance availability [@liu_health_2022] and many others.
 
@@
 ## Summary Statistics
+This section provides an overview of household income and wealth in China from 2011 to 2019, demonstrating that the 2017 and 2019 waves of the CHFS align with earlier studies in confirming a declining trend in China's inequality since 2010.
+
 ::: {#tbl-tbl1}
@@
 
-All figures are reported in RMB at constant 2017 prices.
+*All figures are reported in RMB at constant 2017 prices.*
 
-Sources: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.
+*Sources: @tan_earnings_2017's calculation based on the 2010 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.*
 
@@
 
-Compared to the 2011 data, the Gini coefficient for income has decresed from 0.664 in 2011 to 0.605 in 2017 and remained stable in 2019.
-The Gini coefficient for wealth has decreased from 0.761 in 2011 to 0.701 in 2017 and continued to decrease to 0.667 in 2019.
-The results show that China's income and wealth inequality has been decreasing since 2011.
-The finding is consistent with the findings in @kanbur_great_2021 and @zhang_survey_2021, that China's inequality peaked around 2010 and started to decrease since then.
+Compared to 2010 data, the Gini coefficient for income decreased from 0.664 in 2011 to 0.605 in 2017 and remained stable in 2019.
+Similarly, the Gini coefficient for wealth declined from 0.761 in 2011 to 0.701 in 2017, and continued to decrease to 0.667 in 2019.
+These results indicate a consistent reduction in China's income and wealth inequality since 2010, aligning with findings from @kanbur_great_2021 and @zhang_survey_2021, which also report that China's inequality peaked around 2010 before starting to decline.
 
-Notably, while the growth rate of mean and median income from 2011 to 2017 substantially outpaced that of wealth, this pattern reversed between 2017 and 2019, where the growth of mean and median wealth greatly surpassed income growth.
-This later-period divergence may reflect factors such as significant asset price appreciation relative to income gains.
+Notably, while the growth rate of mean and median income from 2010 to 2017 substantially outpaced that of wealth, this pattern reversed between 2017 and 2019, when the growth in mean and median wealth considerably surpassed that of income.
+This later-period divergence may reflect factors such as significant asset price appreciation relative to income growth.
 
 ## Factor decomposition of wealth inequality
-
-To understand the sources driving wealth inequality in China, this section applies the decomposition method developed by @lerman_income_1985.
-According to this method, the contribution of each wealth component to the overall Gini coefficient is decomposed as follows:
+This section employs the decomposition method developed by @lerman_income_1985 to identify the specific contributions of different asset types to overall wealth inequality in China.
+The decomposition facilitates a detailed examination of how various asset components individually contribute to the aggregate wealth inequality measured by the Gini coefficient.
+This approach enables an analysis of the structural composition of inequality, clarifying not only the relative magnitude of each asset type's impact but also the extent to which each correlates with the overall wealth distribution.
+According to this method, the overall wealth Gini coefficient can be decomposed into contributions from each asset component as follows:
 
@@
 where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth.
-This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking [@lerman_income_1985].
+This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking.
+The summary statistics of each asset type are presented in @tbl-wealth-summary.
 
@@
 
-Source: Author's calculation based on the 2017 and 2019 CHFS data.
+*Source: Author's calculation based on the 2017 and 2019 CHFS data.*
 
-Summary Statistics of Wealth, 2017 and 2019
+*Summary Statistics of Wealth, 2017 and 2019*
 :::
 
-@tbl-wealth-summary summarizes household wealth allocation in China for 2017 and 2019.
-Housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types.
+The data shows that housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types.
 Although housing significantly contributes to overall wealth inequality, homeownership remains widespread, with only about 10% of households reporting no housing assets.
@@
 Land asset ownership, reported by about 44% of households in 2017 and 41% in 2019, closely aligns with China's rural population proportion of approximately 42% in 2017 and 40% in 2019 [@world_bank_rural_nodate].
-However, later results in this paper show that the gini coefficient for land in the rural area was 89.7% in the rural area in 2019, which is significantly higher than the gini coefficient for wealth and for housing in rural areas.
+However, later results in this paper show that the gini coefficient for land in the rural area was as high as 0.9 in the rural area in 2019, which is significantly higher than the gini coefficient for wealth and for housing in rural areas.
 This shows that dispite a wide distribution of land ownership, the distribution of land is highly unequal in modern China.
@@
 
-Source: Author's calculation based on the 2017 CHFS data.
+*Source: Author's calculation based on the 2017 CHFS data.*
 
-Factor Decomposition of Wealth Gini Coefficient, 2017
+*Factor Decomposition of Wealth Gini Coefficient, 2017*
 :::
@@
 
-Source: Author's calculation based on the 2019 CHFS data.
+*Source: Author's calculation based on the 2019 CHFS data.*
 
-Factor Decomposition of Wealth Gini Coefficient, 2019
+*Factor Decomposition of Wealth Gini Coefficient, 2019*
 :::
 
-@tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019 show the factor decomposition results for 2017 and 2019, respectively.
-Both years confirm housing assets' persistent and dominant role in driving China's wealth inequality, contributing over 72% to the overall inequality (72.7% in 2017, 72.3% in 2019).
-This aligns with earlier studies: @knight_why_2022 reported that housing equity's contribution rose significantly from 64% in 2002 to 79% in 2013, while @tan_earnings_2017 documented housing's share at approximately 67% of household wealth in 2011.
-Financial equity emerged as another important factor, increasing its contribution from approximately 11.5% in 2017 to about 13.5% in 2019.
-In contrast, contributions from commercial assets declined notably, potentially due to definitional changes in the survey.
+The factor decomposition results for 2017 and 2019 are presented in @tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019, respectively.
+Both years highlight the persistent and dominant role of housing assets in driving wealth inequality in China, contributing over 72% to overall inequality.
+Specifically, housing assets accounted for 74.2% of total household wealth in 2017 and 73% in 2019, figures notably higher than the 67% reported for 2010 by @tan_earnings_2017 using the same CHFS data.
+Together with the finding by @knight_why_2022, who documented a substantial increase in housing equity’s contribution from 64% in 2002 to 79% in 2013 based on CHIP data, these results confirm that housing has become an increasingly significant factor in shaping wealth inequality in China during the 21st century.
+
+Financial equity represented another significant factor, with its contribution to overall inequality increasing from approximately 11.2% in 2017 to 13.5% in 2019.
+Despite its comparatively smaller share of total household wealth, which limits its overall contribution to inequality, financial equity demonstrates a higher degree of internal inequality in its distribution than housing assets.
+Commercial assets, encompassing agricultural assets, business assets, and storefronts, ranked as the third-largest contributor to overall inequality.
+This asset class exhibits a very high Gini coefficient, potentially attributable to the concentration of family farming and business operations among a relatively small number of households.
+Land and vehicles represented smaller contributions to overall inequality.
+Sharp fluctuations observed in the share and contribution of commercial assets and land to the overall Gini coefficient appear to stem from definitional changes or systemic asset re-evaluation within the survey methodology, warranting caution in interpreting results related to these components across survey waves.
+Debt generally acts as a negative contributor to overall inequality, suggesting that higher-wealth households tend to hold larger absolute amounts of debt.
+An exception is noted for other debt, including medical, credit card, education, and miscellaneous debts, which show a slight positive contribution to overall inequality.
+This indicates that lower-wealth households are more likely to be burdened by consumer debt compared to debt acquired for asset accumulation.
 
 Overall, the structure of household wealth inequality in China remained relatively stable between 2017 and 2019.
-Housing assets continued to play a dominant role, with financial assets becoming increasingly significant, suggesting evolving household investment behaviors.
-To explore heterogeneity in these patterns, the following sections decompose wealth inequality within specific demographic subgroups.
+Housing assets maintained their dominant role, while financial assets became increasingly significant.
+To explore heterogeneity in wealth inequality patterns, the following sections decompose wealth inequality within specific demographic subgroups.
 

### 2025-04-26 report/draft.qmd
update

@@
 
-*Summary Statistics of Wealth, 2017 and 2019*
+Summary Statistics of Wealth, 2017 and 2019
 :::
@@
 
-*Factor Decomposition of Wealth Gini Coefficient, 2017*
+Factor Decomposition of Wealth Gini Coefficient, 2017
 :::
@@
 
-*Factor Decomposition of Wealth Gini Coefficient, 2019*
+Factor Decomposition of Wealth Gini Coefficient, 2019
 :::
@@
 
-Financial equity represented another significant factor, with its contribution to overall inequality increasing from approximately 11.2% in 2017 to 13.5% in 2019.
+Financial equity, including deposit, equity, bond and other financial assets, represented another significant factor, with its contribution to overall inequality increasing from approximately 11.2% in 2017 to 13.5% in 2019.
 Despite its comparatively smaller share of total household wealth, which limits its overall contribution to inequality, financial equity demonstrates a higher degree of internal inequality in its distribution than housing assets.
@@
 ## Factor decomposition partitioned by rural and urban residence
-Since changes in the factor decomposition from 2017 to 2019 for rural and urban households closely mirror those observed for the entire population, only the 2019 decomposition results for rural and urban groups are presented here.
+Since changes in the factor decomposition for rural and urban households from 2017 to 2019 closely resemble those observed for the total population, this study presents only the 2019 decomposition results for these subgroups.
+As the subgroup decompositions utilize data from the same wave of the CHFS, the results remain strictly comparable.
 
@@
 
-Source: Author's calculation based on the 2019 CHFS data.
+*Source: Author's calculation based on the 2019 CHFS data.*
 
@@
 
-The comparative factor decompositions for 2019 underscore several salient distinctions between the determinants of wealth inequality in urban and rural China.
-In both settings residential property remains the predominant source of inequality; nevertheless, its influence is markedly stronger in cities, where housing contributes roughly three‑quarters of the overall Gini coefficient, compared with just over one‑half in the countryside.
-This pronounced urban effect reflects the exceptionally high and spatially differentiated prices of metropolitan real estate as well as the tendency for affluent urban households to own multiple dwellings.
+The comparative factor decompositions for 2019 highlight several notable distinctions between the determinants of wealth inequality in urban and rural China.
+In both contexts, residential property remains the dominant contributor to inequality; however, its impact is significantly more pronounced in urban areas, where housing accounts for approximately three-quarters of the overall Gini coefficient, compared to slightly over half in rural regions.
+This heightened urban influence reflects the disproportionately large share of housing assets within urban households' wealth portfolios.
+Although housing contributes less to inequality and homeownership rates are higher in rural areas (95% compared to 87% in urban areas according to the 2019 CHFS data), the value distribution of housing assets is more unequal in rural areas, as indicated by a higher Gini coefficient (0.67 in rural areas compared to 0.64 in urban areas).
 
-By contrast, assets that are either absent from, or less significant in, urban portfolios assume a more prominent role in rural areas.
-Most notably, agricultural and homestead land accounts for almost one‑fifth of rural inequality but for only about six per cent of the urban total.
-The unequal distribution of land holdings, combined with large inter‑village variation in implicit land values, therefore constitutes a key driver of rural wealth dispersion.
-Commercial (business) assets exhibit a similar pattern: they explain roughly eleven per cent of rural inequality but just seven per cent in cities, mirroring the heterogeneity of small‑scale entrepreneurial activities in village and township economies.
-Conversely, financial equity is slightly more important in urban areas (around thirteen per cent) than in rural areas (about eleven per cent), consistent with better access of urban households to formal financial markets.
+Land, a relatively insignificant asset class in urban household portfolios contributing only 5.8% to urban wealth inequality, represents a far more significant asset for rural households.
+Specifically, land comprises 17.1% of total rural assets and accounts for 18.4% of rural inequality.
+Similar to housing, the ownership rate of land among rural households is high, reaching 81% in 2019.
+Nevertheless, the distribution of land assets exhibits substantial inequality, with a Gini coefficient of 0.9 in 2019, considerably higher than the corresponding coefficient for housing in rural areas (0.67).
 
-Finally, debt liabilities play a modest equalising role in both residence types, as indicated by their negative contributions to the Gini.
-The mitigating effect is somewhat larger in urban China: mortgage debt alone reduces the urban coefficient by roughly 3.3 percentage points, relative to 2.1 percentage points in rural areas.
-Taken together, these findings imply that policy initiatives aimed at curbing wealth inequality ought to be tailored to local contexts.
-In cities, measures that temper speculative housing demand and widen access to diversified financial products are likely to prove most effective, whereas in rural regions reforms that strengthen land‑tenure security, facilitate land transactions, and support productive agricultural and non‑farm enterprises are likely to yield greater distributive gains.
+Financial equity constitutes a greater proportion of wealth and contributes more substantially to wealth inequality in urban areas than in rural areas.
+Nevertheless, the distribution of financial equity remains more unequal among rural households.
+Conversely, vehicle assets account for a higher share of total wealth and contribute more significantly to inequality in rural regions.
+These outcomes regarding financial equity and vehicle assets align with expectations, as rural households are, on average, considerably poorer than urban households, resulting in higher thresholds for acquiring substantial financial assets and vehicles in rural settings.
+Commercial assets also represent a larger share of both total wealth and inequality in rural areas, possibly reflecting the prevalence of family farming and small-scale businesses among rural households.
 
-When benchmarked against the national decompositions for 2017 and 2019 (see @tbl-gini-decomposition-2017 and @tbl-gini-decomposition-2019), the urban–rural patterns largely reiterate the centrality of housing, land, and financial equity for Chinese wealth inequality.
-What the residence‑specific results add, however, is a clearer sense of _relative_ importance: housing is even more dominant in cities than the country‑wide average, whereas land and business assets loom larger in rural portfolios than the aggregate figures would suggest.
-At the same time, the contribution of financial assets in both residence types remains close to the national share, confirming that the rise of household financial investment is a broad‑based rather than location‑specific phenomenon.
+In conclusion, this section highlights the pronounced differences between urban and rural wealth inequality in China as of 2019.
+Urban wealth inequality is driven primarily by the concentration of housing and financial assets, whereas rural wealth inequality is more substantially influenced by disparities in land holdings, commercial assets, and vehicle ownership.
 
 ## Factor decomposition partitioned by east and non-east regions
+This section conducts a factor decomposition analysis for the East and Non-East regions to examine regional differences in wealth inequality.
+The East region predominantly comprises coastal areas, which benefit from proximity to developed East Asian markets, including Japan, South Korea, Taiwan, and Hong Kong.
+In contrast, the Non-East region consists mainly of inland areas, encompassing the central, northwest, and southwest parts of the country, which generally exhibit lower levels of economic development relative to the East.
+
+::: {#fig-map}
+
+![](map.jpg){width=50%, fig-align=center}
+
+*Source: @map*
+
+China Four Regions Map
+:::
+
 It is worth noting that from 2017 to 2019, housing's contribution to the Gini coefficient rose by 2.4 percentage points among households in the East region, whereas it decreased by 2.9 percentage points among Non-East households, despite housing's overall contribution remaining nearly identical across both years.
@@
 
-Relative to the country‑wide decompositions for 2017 and 2019, the regional results confirm the pervasive importance of housing, land and financial assets in shaping wealth inequality across China, yet they also reveal distinct regional configurations.
-In the coastal East, housing's influence is particularly pronounced; its contribution climbs to over 80 per cent (80.3%) of total inequality—roughly eight percentage points above the national average—underscoring the outsized role of metropolitan real‑estate markets and its strong alignment with overall wealth rank (Gini correlation 0.96).
-By contrast, in the Non‑East provinces housing accounts for a lower share at 58.3 per cent.
-Here, land and commercial (business) assets play a much larger role, together explaining nearly one-quarter of inequality (23.2%, combining 13.8% from land and 9.4% from commercial), substantially higher than their combined share of about 7.6% in the East.
-This reflects not only portfolio differences but also stronger correlations between these assets and wealth rank in the Non-East (Gini correlations of 0.83 for land and 0.80 for commercial, vs.
-0.35 and 0.72 respectively in the East).
+The differences observed in the factor decomposition between the East and Non-East regions closely resemble those identified between urban and rural areas.
+Part of this similarity can likely be attributed to the higher urbanization rate in the East region compared to the Non-East regions (71% versus 61% in the 2019 CHFS data).
 
-Financial equity is a sizeable driver in both regions, though its contribution and correlation with wealth rank are marginally greater in the interior (15.7% contribution, 0.78 correlation) than along the coast (12.4% contribution, 0.77 correlation), perhaps reflecting the recent expansion of retail investment beyond major coastal cities.
-Debt components, particularly mortgage liabilities, serve to dampen inequality in both regions through their negative contributions to the overall Gini coefficient.
-These contrasts underscore that policies aimed at moderating wealth disparities need to account for regional heterogeneity—potentially addressing speculative housing dynamics in the East while simultaneously fostering land‑market reforms and business development in the interior—within a common framework of wider financial inclusion.
+Relative to the nationwide decompositions for 2017 and 2019, the regional analyses reaffirm the pervasive significance of housing, land, and financial assets in shaping wealth inequality across China, while simultaneously highlighting distinctive regional patterns.
+In the coastal East, the influence of housing is particularly pronounced, contributing over 80 percent (80.3%) of total inequality—approximately eight percentage points above the national average.
+In contrast, housing accounts for a comparatively smaller share (58.3%) in the Non-East regions.
+Across all three decomposition components, housing distribution in the East aligns more closely with overall wealth rankings, is more unequal, and constitutes a greater share of total household wealth.
+
+Land represents a minor asset class in the East region, accounting for only 2.6% of total household wealth, and displays a weak correlation with overall wealth rankings.
+Conversely, similar to the urban-rural comparison, land significantly influences wealth inequality in the inland regions, comprising 11.1% of total wealth and contributing 13.8% to total inequality.
+Unlike in the urban-rural comparison, financial equity emerges as a more critical asset class and driver of inequality in the Non-East regions than in the East.
+Specifically, financial equity constitutes 17.2% of total wealth and 15.7% of total inequality in the Non-East, compared to just 14.1% and 12.4%, respectively, in the East.
+The distribution patterns for commercial and vehicle assets closely align with those observed in the urban-rural comparisons.
+
+In conclusion, while housing remains the dominant contributor to wealth inequality in both East and Non-East regions, land and financial equity play relatively more significant roles in shaping inequality in the inland regions compared to the coastal East.
+Notably, financial equity is a stronger driver of inequality in the Non-East regions than in the East, a somewhat unexpected finding that may be explained by the comparatively lower housing values in inland areas, prompting households there to allocate a greater proportion of their wealth to financial assets.
 
 ## Group decomposition of wealth inequality by rural and urban
-To analyze the sources of wealth inequality, this paper employs the classic Gini decomposition identity initially developed by Bhattacharya and Mahalanobis (1967)[@bhattacharya_regional_1967].
+After examining both the overall landscape and the potential heterogeneity of wealth inequality in China, this section further investigates the extent to which national wealth inequality can be attributed to urban-rural and regional disparities, as well as to inequality within these subgroups.
+To analyze the sources of wealth inequality, this paper employs the classic Gini decomposition identity initially developed by @bhattacharya_regional_1967.
 The overall Gini coefficient  is decomposed as follows:
@@
 where $G_{B}$ is the between-group Gini, which captures inequality arising solely from differences in group means; $G_{k}$ is the within-group component, a weighted sum of the inequality within each group, with weights $a_{k}$ representing the products of each group's population share and wealth share; and $R$ is the overlap (residual) component, which captures inequality due to the intersection of group wealth distributions.
-Lambert and Aronson (1993) provide a geometric interpretation of this overlap term, demonstrating that $R$ corresponds to the Lorenz curve area generated by reranking when subgroup wealth distributions overlap.
-A positive $R$ thus indicates significant intersections among wealth distributions of different groups[@lambert_inequality_1993].
+@lambert_inequality_1993 provide a geometric interpretation of this overlap term, demonstrating that $R$ corresponds to the Lorenz curve area generated by reranking when subgroup wealth distributions overlap.
+A positive $R$ thus indicates significant intersections among wealth distributions of different groups.
 
@@
 
-Source: Author's calculation based on the 2017 and 2019 CHFS data.
+*Source: Author's calculation based on the 2017 and 2019 CHFS data.*
 
@@
 
-The decomposition results indicate that about 55–57% (55.4% in 2017, 56.8% in 2019) of total wealth inequality is attributable to within-group disparities among urban and rural households.
-The between-group component, reflecting the urban-rural wealth divide, contributes approximately 33–34% (32.6% in 2017, 33.7% in 2019).
-The overlap component, representing inequality due to intersecting distributions of urban and rural wealth, accounts for around 9–12% (12.0% in 2017, 9.5% in 2019) of total inequality.
-The consistency of these shares between 2017 and 2019 underscores the continued significance of the urban-rural divide in shaping China's wealth inequality.
+The decomposition results indicate that approximately 55–57% (55.4% in 2017, 56.8% in 2019) of total wealth inequality can be attributed to disparities within urban and rural households.
+Meanwhile, the between-group component, reflecting differences in average wealth between urban and rural households, contributes roughly 33–34% (32.6% in 2017, 33.7% in 2019).
+Additionally, the overlap component, capturing inequality arising from intersecting wealth distributions of urban and rural groups, accounts for about 9–12% (12.0% in 2017, 9.5% in 2019) of overall inequality.
+The stability of these shares from 2017 to 2019 highlights the persistent importance of both within-group inequalities and the urban-rural wealth divide in shaping China’s national wealth inequality.
 
@@
 
-The group decomposition shows that, similar to the urban-rural decomposition, the between-group component contributes over 30% to overall wealth inequality in both 2017 and 2019 (36.9% and 33.8% respectively).
-The between-group contribution saw a slight decrease from 2017 to 2019, with a similar increase in the residual component (from 32.0% to 34.5%).
-This might signal some modest convergence in household wealth between regions.
+The group decomposition indicates that, analogous to the urban-rural analysis, the between-group component accounted for over 30% of total wealth inequality in both 2017 and 2019 (36.9% and 33.8%, respectively).
+Between-group contributions experienced a slight decline from 2017 to 2019, mirrored by a corresponding increase in the residual component (from 32.0% to 34.5%).
+This shift may suggest a modest convergence in household wealth across regions.
+Nonetheless, it remains clear that China's wealth inequality arises significantly from both within-region disparities and regional differences, with each component consistently contributing over 30% to overall inequality.
 
-## Conclusion
-This paper analyzed wealth inequality in China using the 2017 and 2019 CHFS data, finding a modest decline in the overall Gini coefficient (0.701 to 0.667), consistent with literature suggesting a post-2010 peak [@kanbur_great_2021; @zhang_survey_2021].
-Factor decomposition confirmed housing assets as the primary driver (contributing over 72%), aligning with previous studies [@knight_why_2022; @tan_earnings_2017], while financial equity's contribution grew (from approximately 11% to 14%).
-Partitioned analysis highlighted housing's stronger impact in urban areas (around 75% vs.
-approximately 52% rural) and especially the East region (over 80% vs.
-approximately 58% Non-East), with divergent trends emerging between 2017-2019 (increasing contribution in the East, decreasing elsewhere).
-Group decomposition further revealed that while within-group inequality dominates (accounting for approximately 55-60%), substantial inequality stems from persistent urban-rural (around 33%) and regional (over 30%) divides.
+# Conclusion and policy implications
+This paper analyzed wealth inequality in China using data from the 2017 and 2019 China Household Finance Survey (CHFS).
+The findings confirm arguments from earlier studies on China's income inequality, indicating that the country's wealth inequality has plateaued since 2010.
+Specifically, the national wealth Gini coefficient declined from 0.76 in 2010 to 0.70 in 2017, and further decreased to 0.67 in 2019.
 
-These dynamics likely reflect intertwined factors: rapid urbanization, uneven real estate appreciation and speculation impacting housing's role; financial market development driving the rise of financial assets, potentially favoring wealthier urban households; and deep-rooted structural issues like historical policies, the Hukou system, and differential access to opportunities maintaining the significant urban-rural and regional gaps, despite potential slow convergence hinted by recent trends and policies [@kanbur_great_2021].
-Consequently, addressing China's high wealth inequality requires a nuanced approach beyond national measures, necessitating context-specific policies such as targeted housing market regulation in the East, support for land and business development in rural and Non-East areas, and broader efforts towards financial inclusion across the country.
+Employing detailed factor and group decomposition analyses, this study demonstrated that wealth inequality in China is a multifaceted phenomenon, driven simultaneously by urban-rural disparities, regional differences, and within-group inequality.
+Moreover, substantial heterogeneity was observed in the sources and extent of wealth inequality across various subgroups, particularly between rural and urban households, as well as between the East and Non-East regions.
+
+These findings carry significant policy implications.
+To effectively reduce China's wealth inequality, policymakers must address both between-group disparities, such as those between urban-rural and regional populations, and within-group inequalities.
+Addressing between-group inequality necessitates targeted policies specifically aimed at improving conditions in rural and Non-East regions.
+Meanwhile, policies to mitigate within-group inequality should carefully consider the heterogenous characteristics of wealth distribution in different subgroups.
+For instance, specialized interventions are required in urban and Eastern regions to reduce housing-related inequality, whereas in rural and Non-East regions, measures such as equitable land redistribution and increased support for commercial and vehicle asset accumulation may prove more effective.
 

### 2025-05-16 report/Chenning_paper.qmd
update

--- 2025-04-26 report/draft.qmd
@@
         linestretch:  1.2
-        fontsize: 10pt
+        fontsize: 12pt
+        author: "Chenning Xu"
         fig-pos: "H"
-        author: "Chenning Xu"
 bibliography: tom.bib
+abstract: "This paper analyzes recent trends and structural determinants of wealth inequality in China using microdata from the 2017 and 2019 China Household Finance Survey (CHFS).
+The findings indicate a moderation in wealth inequality, aligning with earlier research focusing primarily on income inequality.
+By employing factor decomposition and group decomposition techniques, the analysis highlights considerable heterogeneity in the profiles of wealth inequality between urban-rural and coastal-inland populations.
+It further demonstrates that both urban-rural and regional disparities substantially contribute to overall wealth inequality.
+Based on these findings, the paper argues that effectively addressing wealth inequality in China requires government policies targeting urban-rural and regional imbalances, while also addressing within-group inequalities according to their distinct inequality profiles."
 ---
@@
 Coastal provinces, for instance, often experienced higher growth rates, benefiting from earlier integration with global markets, while inland and rural areas developed more slowly, contributing to regional disparities [@liao_sixty_2016].
-Sectorally, growth was initially propelled by fixed-capital investment and export-oriented manufacturing, a model that later showed signs of creating imbalances [@world_bank_world_2024].
+Sectorally, growth was initially propelled by fixed-capital investment, residential investment boom, andexport-oriented manufacturing, a model that later showed signs of creating imbalances [@world_bank_world_2024].
 China's annual GDP growth rate moderated from 10.6% in 2010 to 5.2% by 2023, reflecting a slowdown from the high rates experienced earlier in the decade [@world_bank_gdp_2025].
@@
 
-This study contributes to the understanding of China's wealth inequality by employing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS).
-The analysis focuses specifically on assessing whether wealth inequality trends during this later period substantiate claims of ongoing inequality reduction observed in earlier research focused primarily on income [@kanbur_great_2021; @zhang_survey_2021].
-The paper utilizes factor decomposition methods to pinpoint the contributions of various asset types to overall wealth inequality.
-A further level of analysis involves applying this factor decomposition separately to distinct population subgroups defined by urban-rural residence and coastal-inland location, allowing for an examination of potential heterogeneity in the drivers of inequality across these dimensions.
-Additionally, group decomposition techniques are used to quantify the portion of total wealth inequality attributable to the average wealth differences between these urban-rural and coastal-inland groups.
-This combination of recent data and multiple decomposition strategies aims to offer a detailed perspective on the patterns and structural underpinnings of wealth inequality in contemporary China.
+This study advances the understanding of wealth inequality in China by analyzing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS).
+Specifically, it evaluates whether recent wealth inequality trends align with earlier findings of declining inequality, which were predominantly based on income data [@kanbur_great_2021; @zhang_survey_2021].
+The paper employs factor decomposition to identify the contributions of different asset types to overall wealth inequality.
+This approach is also applied separately to population subgroups defined by urban-rural and coastal-inland divisions to explore potential heterogeneity in the determinants of inequality across these dimensions.
+Additionally, group decomposition is used to quantify how much of total wealth inequality can be attributed to differences in average wealth between and within these subgroups.
+By combining recent data with these decomposition techniques, the paper provides an in-depth examination of the patterns and structural determinants underpinning wealth inequality in contemporary China.
+Based on these findings, this paper argues that government policies aiming to reduce China’s wealth inequality should simultaneously target urban-rural and regional imbalances, as well as address inequality within each subgroup according to their specific inequality profiles.
 
@@
 
-The CHFS provides sample weights, calculated based on the multi-stage sampling design and adjusted using population benchmarks, to facilitate representative analysis.
 The dataset has been utilized in numerous academic studies exploring various facets of the Chinese economy and household behavior.
@@
 This section provides an overview of household income and wealth in China from 2011 to 2019, demonstrating that the 2017 and 2019 waves of the CHFS align with earlier studies in confirming a declining trend in China's inequality since 2010.
+
+Compared to 2010 data, the Gini coefficient for income decreased from 0.664 in 2011 to 0.605 in 2017 and remained stable in 2019.
+Similarly, the Gini coefficient for wealth declined from 0.761 in 2011 to 0.701 in 2017, and continued to decrease to 0.667 in 2019.
+These results indicate a consistent reduction in China's income and wealth inequality since 2010, aligning with findings from @kanbur_great_2021 and @zhang_survey_2021, which also report that China's inequality peaked around 2010 before starting to decline.
 
@@
 
+Regarding the causes for post-2000s decline in China’s inequality, @kanbur_great_2021 and @zhang_survey_2021 point to both the Lewis Turning Point and targeted policies.
+By the late 2000s, the absorption of surplus rural labor had tightened rural labor markets, narrowing the rural–urban wage gap.
+At the same time, the 2004 Minimum Wage Law, backed by increasingly rigorous enforcement, and the rollout of rural cooperative medical insurance in 2004, followed by rural social security schemes in 2009, together helped compress both income and wealth disparities.
+
 ## Factor decomposition of wealth inequality
-This section employs the decomposition method developed by @lerman_income_1985 to identify the specific contributions of different asset types to overall wealth inequality in China.
-The decomposition facilitates a detailed examination of how various asset components individually contribute to the aggregate wealth inequality measured by the Gini coefficient.
-This approach enables an analysis of the structural composition of inequality, clarifying not only the relative magnitude of each asset type's impact but also the extent to which each correlates with the overall wealth distribution.
-According to this method, the overall wealth Gini coefficient can be decomposed into contributions from each asset component as follows:
+This section applies the decomposition method developed by @lerman_income_1985 to identify the distinct contributions of various asset types to overall wealth inequality in China.
+This decomposition provides a detailed examination of how each asset component influences aggregate wealth inequality as measured by the Gini coefficient.
+Specifically, it separates each asset's contribution into three components: its share in total household wealth, its own Gini coefficient, and its correlation with the overall wealth rank.
+Formally, the overall wealth Gini coefficient can be decomposed into contributions from individual asset components as follows:
 
@@
 where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth.
-This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking.
 The summary statistics of each asset type are presented in @tbl-wealth-summary.
+
+The data shows that housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types.
+Although housing significantly contributes to overall wealth inequality, homeownership remains widespread, with only about 10% of households reporting no housing assets.
+Moreover, housing-related debt remains minimal relative to housing values, with approximately 82% of households reporting no such debts, suggesting broad equity-based homeownership patterns.
 
@@
 :::
-
-The data shows that housing assets continue to dominate household wealth portfolios, consistently demonstrating much higher mean and median values compared to other asset types.
-Although housing significantly contributes to overall wealth inequality, homeownership remains widespread, with only about 10% of households reporting no housing assets.
-Moreover, housing-related debt remains minimal relative to housing values, with approximately 82% of households reporting no such debts, suggesting broad equity-based homeownership patterns.
 
@@
 
+The differences observed in the factor decomposition between the East and Non-East regions closely resemble those identified between urban and rural areas.
+Part of this similarity can likely be attributed to the higher urbanization rate in the East region compared to the Non-East regions (71% versus 61% in the 2019 CHFS data).
+
 ::: {#tbl-east-decom}
@@
 
-The differences observed in the factor decomposition between the East and Non-East regions closely resemble those identified between urban and rural areas.
-Part of this similarity can likely be attributed to the higher urbanization rate in the East region compared to the Non-East regions (71% versus 61% in the 2019 CHFS data).
-
 Relative to the nationwide decompositions for 2017 and 2019, the regional analyses reaffirm the pervasive significance of housing, land, and financial assets in shaping wealth inequality across China, while simultaneously highlighting distinctive regional patterns.
-In the coastal East, the influence of housing is particularly pronounced, contributing over 80 percent (80.3%) of total inequality—approximately eight percentage points above the national average.
+In the coastal East, the influence of housing is particularly pronounced, contributing over 80 percent (80.3%) of total inequality, approximately eight percentage points above the national average.
+It is even fair to say that "housing inequality is the wealth inequality" in the East region.
 In contrast, housing accounts for a comparatively smaller share (58.3%) in the Non-East regions.
@@
 
-The above summary statistics shows that the mean and median wealth in the East is significantly higher than in the other regions in both 2017 and 2019.
-Notably, the mean wealth of West-area households surpassed that of the Middle region in 2019, although its median wealth remained lower.
-Concurrently, the West's Gini coefficient rose to become the highest among all regions in 2019 (0.6947).
+The summary statistics above indicate that both the mean and median household wealth in the East were significantly higher than those in the other regions in 2017 and 2019.
+Notably, in 2019, the mean wealth of households in the Western region surpassed that of the Middle region, although its median wealth remained lower.
+Concurrently, the Gini coefficient for the West increased sharply, becoming the highest among all regions in 2019 (0.6947).
+However, there is a lack of empirical literature examining the underlying causes for the rise in wealth inequality observed in the Western region during this period.
 
@@
 
-The group decomposition indicates that, analogous to the urban-rural analysis, the between-group component accounted for over 30% of total wealth inequality in both 2017 and 2019 (36.9% and 33.8%, respectively).
+The regional group decomposition indicates that, analogous to the urban-rural analysis, the between-group component accounted for over 30% of total wealth inequality in both 2017 and 2019 (36.9% and 33.8%, respectively).
 Between-group contributions experienced a slight decline from 2017 to 2019, mirrored by a corresponding increase in the residual component (from 32.0% to 34.5%).
@@
 
+In conclusion, the group decomposition results indicate that structural imbalances in China, primarily those between urban and rural areas and between East and Non-East regions, significantly contribute to overall wealth inequality.
+Each of these between-group effects individually accounts for over one-third of the total wealth inequality in 2019.
+Given that the residual component in each decomposition captures portions of between-group effects from the alternative grouping, the combined contribution of these structural imbalances likely surpasses one-third by some notable percentage points.
+Furthermore, the within-group contribution in the urban-rural decomposition is substantial, reaching 56.8% in 2019.
+Although the within-group contribution in the regional decomposition is relatively lower (31.5%), the combined effect of within-group inequality and residual overlap (34.5%) remains considerable.
+Consequently, policy efforts must not only focus on addressing structural disparities between different demographic groups but also actively mitigate inequalities within each population subgroup.
+
 # Conclusion and policy implications
 This paper analyzed wealth inequality in China using data from the 2017 and 2019 China Household Finance Survey (CHFS).
-The findings confirm arguments from earlier studies on China's income inequality, indicating that the country's wealth inequality has plateaued since 2010.
-Specifically, the national wealth Gini coefficient declined from 0.76 in 2010 to 0.70 in 2017, and further decreased to 0.67 in 2019.
+The findings substantiate earlier studies focused on China's income inequality, suggesting that the country's wealth inequality has stabilized since 2010.
+Specifically, the national wealth Gini coefficient decreased from 0.76 in 2010 to 0.70 in 2017, and further declined to 0.67 in 2019.
+A comprehensive understanding of the drivers behind this moderation in inequality requires more rigorous causal analyses covering this recent period.
 
